### 2018-09-20 manuscript/eq_cycle_slip_rates.tex
latex, figs and new results

\documentclass[se, manuscript]{copernicus}
\usepackage{mathptmx}

\begin{document}

\title{The impact of earthquake cycle variability on neotectonic and
paleoseismic slip rate estimates}

\Author[1, 2, 3]{Richard}{Styron}

\affil[1]{Earth Analysis, 21855 Bear Creek Road, Los Gatos, CA 95033 USA}
\affil[2]{Global Earthquake Model Foundation, Via Ferrata 1, Pavia 27100 Italy}
\affil[3]{Department of Geology, University of Kansas, Ritchie Hall, Earth Energy \& Environment Center, 1414 Naismith Drive,
Room 254, Lawrence, KS 66054 USA}

\runningtitle{Impact of earthquake cycle on slip rates}
\runningauthor{Richard Styron}
\correspondence{Richard Styron (richard.h.styron@gmail.com)}

\firstpage{1}
\maketitle

\begin{abstract}
Because of the natural variability (aleatoric uncertainty) in earthquake
recurrence intervals and coseismic displacements on a fault, cumulative
slip on a fault does not increase linearly or perfectly step-wise with
time; instead, some amount of variability in shorter-term slip rates
results. Though this variability could greatly affect the accuracy of
neotectonic (i.e., late Quaternary) and paleoseismic slip rate
estimates, these effects have not been quantified. In this study,
idealized faults with four different, representative earthquake
recurrence distributions are created with equal mean recurrence
intervals (1,000 years) and coseismic slip distributions, and the
variability in slip rate measurements over 500 to 100,000 year
measurement windows is calculated for all faults through Monte Carlo
simulations. The recurrence distributions used are quasi-periodic,
unclustered and clustered lognormal distributions, and an unclustered
exponential distribution. The results demonstrate that the most
important parameter is the coefficient of variation (\(COV\) = standard
deviation / mean) of the recurrence distributions rather than the shape
of the distribution itself. Slip rate variability over short time scales
(\textless{} 10,000 years or 10 mean earthquake cycles) is quite high,
but decreases with time and is close to stable after
\textasciitilde{}40,000 years (40 mean earthquake cycles). This
variability is higher for recurrence distributions with a higher
\(COV\). The natural variability in the slip rate estimates compared to
the true value is then used to estimate the epistemic uncertainty in a
single slip rate measurement (as one would make in a geological study)
in the absence of any measurement uncertainty. This epistemic
uncertainty is very high (a factor of 2 or more) for measurement windows
of a few mean earthquake cycles (as in a paleoseismic slip rate
estimate), but decreases rapidly to a factor of 1-2 with \textgreater{}5
mean earthquake cycles (as in a neotectonic slip rate study). These
uncertainties are independent of, and should be propagated with
uncertainties in fault displacement and geochronologic measurements used
to estimate slip rates. They may then aid in the comparison of slip
rates from different methods or the evaluation of potential slip rate
changes over time.
\end{abstract}

\introduction

Fault slip rates are generally estimated by dividing measurements of the
offset of geologic `marker' features by the time over which that offset
accumulated. The uncertainty in the resulting slip estimate is typically
treated as \emph{epistemic}, and quantified through the propagation of
the measurement uncertainties on the offset and time quantities
\citep[e.g.]{bird_uncertainties_2007,zechar_incorporating_2009}.
However, for slip rate estimates on active faults made from offset
measurements near the fault trace (i.e., within a horizontal distance
that is a small fraction of the fault's locking depth), the episodic
nature of surface displacement due to the earthquake cycle will
necessarily affect the results: If the measurements are taken
immediately before an earthquake, the measured offset and resulting slip
rate estimate will be lower than average, while if the measurements are
taken immediately after an earthquake, the offset and rate will be
higher.

The magnitude of the perturbation to the slip rate is, of course, a
function of the number of cumulative earthquakes that have contributed
to the measured offset. For older Quaternary markers that have
experienced tens to hundreds of major earthquakes, the effects will be
minor, and for bedrock geologic markers with kilometers of displacement,
the earthquake cycle is likely not worth accounting for. However, due to
progressive erosion of geologic markers and the challenge of dating many
late Pliocene to early Quaternary units (which are too old for
radiocarbon and many cosmogenic nuclide systems), geologists often have
no choice but to choose late Pleistocene to Holocene markers to date,
and these units may also be more desirable targets if the scientists are
primarily concerned with estimating the contemporary slip rate on a
fault with a slip rate that may vary over Quaternary timescales
\citep[e.g.]{rittase_temporal_2014,zinke_highly_2018}. For slow-moving
faults, the slip either long waiting to be released, or recently
released, may represent a sizeable fraction of the measured fault
offset.

Careful paleoseismologists and neotectonicists will take this into
account in their slip rate calculations if sufficient data are
available, especially in the years after a major earthquake
\citep[e.g.]{rizza_earthquake_2015}, and many others will discuss the
potential effects if the data are not \citep[e.g.]{lifton_latest_2015}.
These workers may only consider the time since the last earthquake,
often making the assumption (stated or not) that the earthquakes are
identical in slip and perfectly periodic.

However, the recurrence intervals between successive earthquakes on any
given fault segment have some natural variability (i.e. \emph{aleatoric
uncertainty}); similarly, displacement at a measured point is not
identical in each earthquake \citep[e.g.]{duross_holocene_2008}.
Therefore, the measured slip rate may deviate from the time-averaged
rate based on the amount of natural variability in the earthquake cycle,
particularly given successive events from the tails of the recurrence
interval or displacement distributions. Though the physical mechanisms
responsible and the statistical character of this natural variability
remain under debate, its effects on the estimated slip rates may still
be estimated given some common parameterizations.

In this study, the effects of the natural variability in earthquake
recurrence intervals and per-event displacements on neotectonic slip
rate estimates are investigated through Monte Carlo simulations. The
study is geared towards providing useful heuristic bounds on the
aleatoric and epistemic uncertainty of late Quaternary slip rate
estimates for fault geologists, probabilistic seismic hazard modelers,
and others for whom such uncertainties are important.

\section{Modeling the earthquake
cycle}\label{modeling-the-earthquake-cycle}

To study the effects of the natural variability in the earthquake cycle
on estimated slip rates, long displacement histories of a simulated
fault with different parameterizations of the earthquake recurrence
distribution will be created. Then, the mean slip rate over time windows
of various sizes will be calculated from each of the simulated
displacement histories, and the distribution in these results will be
presented, representing the natural variability in this quantity.

To isolate the effects of the earthquake cycle from other phenomena that
may affect slip rate estimates, this study does not attempt to model
erosion, nor does it consider any measurement uncertainty in the age or
offset of the faulted geologic markers; these measurements are assumed
to be perfect. Additionally, though natural variability in
per-earthquake displacement is included in the model, it is minor and
the same for all recurrence distributions. Furthermore, though the model
has one length dimension (fault offset), it is still best thought of as
a point (0-dimensional) model, as there is no spatial reference or
along-strike or down-dip variability, and hence the magnitude of each
earthquake is undefined, and no magnitude-frequency distribution exists.

\subsection{Earthquake recurrence interval
distributions}\label{earthquake-recurrence-interval-distributions}

There are a handful of statistical models for earthquake recurrence
interval distributions that are in widespread consideration.

The most commonly used is the \emph{exponential} distribution. This is
associated with a Poisson process, and is the distribution that results
from a prescribed number of earthquakes being distributed uniformly
randomly within some time interval. Consequently, the probability of an
earthquake (or other event) occurring at any time does not change with
time since the previous event (in other words, the hazard function is
time-invariant); this leads to the characterization of the exponential
recurrence distribution as `random', `memoryless' or `time-independent'.
The exponential distribution is also the simplest to describe
statistically, as it requires only one parameter (the mean rate
parameter), which is a statistical scale parameter. The standard
deviation of a large number of samples generated from an exponential
distribution is equal to the mean.

The other distributions that are in common usage are time-dependent
distributions, meaning that the probability of an event occurring at any
time since the previous event changes with the elapsed time since that
event. This class of distributions includes the lognormal, Weibull, and
Brownian Passage Time \citep{matthews_brownian_2002} distributions.
Though these distributions differ in notable ways, particularly in the
properties of the right tails at several times the mean
\citep{davis_longer_1989,matthews_brownian_2002}, they share a broadly
general shape, and given suitable parameters, generated sample sets of
small size may not be substantively different. In fact, the
distributions are similar enough that it is difficult if not impossible
to discriminate between them given realistic seismologic and
paleoseismologic datasets
\citep{matthews_brownian_2002,ogata_estimating_1999}. These
distributions are described by both the scale and shape parameters.

The behavior of these distributions and of empirical datasets may be
characterized by the regularity of the spacing between events (i.e.~the
recurrence intervals): these may be periodic, unclustered (i.e.,
`random'), or clustered. Assignment into these categories is typically
done with a parameter known as the coefficient of variation, or
\(COV = \sigma / \mu\), where \(\sigma\) is the standard deviation of
the recurrence intervals, and \(\mu\) is the mean recurrence interval.

Periodic earthquakes are those that occur more regularly than random,
and have a \(COV < 1\) (i.e. \(\sigma < \mu\)). These may be generated
by any of the time-dependent distributions described above with suitable
scale and shape parameters. (Note that in this paper, the use of the
term `periodic' does not mean \emph{perfectly} repeating as it might in
the physics or mathematics literature.)

Unclustered earthquakes occur as regularly as random, and have a
\(COV=1\). These may be generated by the exponential distribution (which
can generate no other), or by any of the time-dependent distributions as
well, given the appropriate parameters. Note that sample sets generated
from these different distributions will not be identical, however:
Sequences with an exponential recurrence distribution will have many
more pairs of events that are much more closely spaced together than the
mean, and more pairs of events that are much more widely spaced than the
mean, compared to a sequence generated from the lognormal distribution.
Nonetheless, these will cancel out in the aggregate statistics, so that
the standard deviations will be equal. A comparison of these may be seen
in Figure \ref{eq_rec_dists}.

Clustered earthquake sequences have sets of very tightly spaced
earthquakes that are widely separated, and have a \(COV>1\). These may
be generated from a hyperexponential distribution, which is the sum of
multiple exponentials with different means, or from the time-dependent
distributions given above, given the right parameters.

No consensus exists among earthquake scientists as to the most
appropriate recurrence interval distribution. As is generally the case
with propriety, the safest and probably most correct assumption is that
it is context-dependent. Many studies of plate boundary faults such as
the San Andreas conclude that major or `characteristic' earthquakes are
periodic\\
\citep[e.g.]{berryman_major_2012,scharer_quasi-periodic_2010}.
Conversely, many intraplate faults with low slip rates appear to show
clustered earthquakes separated by long intervals of seismic quiescence
\citep[e.g.]{clark_long-term_2012}. However, one can find examples of
studies indicating the opposite conclusions
\citep{tuttle_earthquake_2002,grant_paleoseismic_1994}.

\subsubsection{Modeled recurrence interval
distributions}\label{modeled-recurrence-interval-distributions}

This study will compare four recurrence interval distributions (Figure
\ref{eq_rec_dists}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  A \emph{periodic} distribution, represented by a lognormal
  distribution with a mean recurrence interval \(\mu\) = 1000 years, and
  a standard deviation \(\sigma\) = 500 years, and a \(COV\) = 0.5.
\item
  An \emph{unclustered} time-dependent distribution, represented by a
  lognormal distribution with a mean recurrence interval \(\mu\) = 1000
  years, a standard deviation \(\sigma\) = 1000 years, and a \(COV\) =
  1.0.
\item
  An \emph{clustered} time-dependent distribution, represented by a
  lognormal distribution with a mean recurrence interval \(\mu\) = 1000
  years, a standard deviation \(\sigma\) = 2000 years, and a \(COV\) =
  2.0.
\item
  An \emph{unclustered} time-independent distribution, represented by an
  exponential distribution with a mean recurrence interval \(\mu\) =
  1000 years, a standard deviation \(\sigma\) = 1000 years, and a
  \(COV\) = 1.0.
\end{enumerate}

These distributions have been selected to represent a diversity of
behaviors with a compact and tractable number of simulations, and
particularly to explore how changes in \(COV\) as well as the shape of
the distribution impact slip rate estimates.

\subsubsection{Earthquake slip
distributions}\label{earthquake-slip-distributions}

All earthquake recurrence distributions share a single earthquake slip
distribution (Figure \ref{slip_dist}). This distribution is a lognormal
distribution with \(\mu\) = 1 m and \(\sigma\) = 0.75 m, which produces
essentially `characteristic' earthquakes that still nonetheless have
some variability. This is representative of behavior observed in many
studies \citep[e.g.]{zielke_slip_2010,klinger_characteristic_2011,zielke_earthquake_2018}. Taken together, the mean slip of 1 m and the
mean recurrence interval of 1000 years shared by each of the recurrence
interval distributions yields a mean slip rate of 1 mm/yr. This rate is
fairly typical for intraplate faults studied by paleoseismologists, and
also allows for easy normalization so that the results of this study can
be generalized to faults with different parameters.

\subsection{Stochastic displacement
histories}\label{stochastic-displacement-histories}

For each of the earthquake recurrence distributions, a 2 million year
long time series of cumulative displacements is calculated, and then
slip rates are estimated over time windows of different lengths.

The construction of the displacement histories is straightforward. From
each recurrence distribution, a little over 2,000 samples are drawn
randomly. Then, these are combined with an equal number of displacement
samples drawn randomly from the earthquake slip distribution. Finally a
cumulative displacement history is created for each series from a
cumulative sum of both the recurrence interval samples (producing an
earthquake time series) and displacement samples (producing a cumulative
slip history). Years with no earthquakes are represented as having no
increase in cumulative displacement. Then, the series is trimmed at
year=2,000,000; it is initially made longer because the stochastic
nature of the sample sets means that 2,000 earthquakes may not always
reach 2,000,000 years.

The displacement histories in Figure \ref{disp_histories} clearly show
that given the stochastic nature of the samples, the cumulative
displacements can diverge greatly from the mean. The magnitude of this
divergence appears to be related to the \(COV\) of the recurrence
interval distributions: The clustered series (\(COV\) = 2) has by far
the most divergence, both unclustered series (lognormal and exponential
with \(COV\) = 1) behave qualitatively similarly, and the periodic
series (\(COV\) = 0.5) tracks most closely with the mean. The
divergences from the mean are driven by successive closely-spaced
earthquakes, perhaps with high displacements, or by long durations of
quiescence. The clustered series in particular shows a pattern of many
closely-spaced events (clusters) leading to a much higher than average
displacement accumulation rate, followed by very long episodes of
dormancy in which regression to the mean occurs. From visual inspection,
the dormant episodes appear to be composed of single or dual
exceptionally long inter-event times. This of course is reflected in the
great asymmetry of this distribution (Figure \ref{eq_rec_dists}), with
the very short mode and `fat' right tail.

Please note that in the construction of the cumulative displacement
histories, all samples are independent. This means that the duration of
any recurrence interval does not depend on the duration of the previous
or subsequent interval (in other words, there is no autocorrelation in
these series); the same applies to the displacement samples.
Furthermore, the magnitude of displacement is independent of the
corresponding recurrence interval. It is currently unknown to what
degree autocorrelation exists in real earthquake time and displacement
series, or how much correlation is present between recurrence intervals
and subsequent displacements. The former is essentially unstudied,
though on the basis of preliminary, unreviewed analysis
\citep{styron_survival_2017}, I suspect that it is as important as
\(COV\). With respect to the latter, the framework of elastic rebound
theory in its most basic form should predict some correspondence between
inter-event (loading) duration and slip magnitude, and this is included
(implicitly or explicitly) in oscillator models incorporating complete
stress or strain release in each earthquake\\
\citep[e.g.]{matthews_brownian_2002,dicaprio_post-seismic_2008} or in
any model where coseismic friction drops to zero, as this is
functionally equivalent (because \(f^a = \tau^a_s / \tau^a_n\), where
\(f^a\), \(\tau^a_s\) and \(\tau^a_n\) are respectively friction at
rupture arrest, shear stress at rupture arrest, and effective normal
stress at rupture arrest, zero friction implies zero shear stress, or
complete stress drop). Given a reasonably constant loading rate,
complete shear stress or strain release implies some proportionality
between the loading time and displacement. Nonetheless, this
correspondence is not found in the more extensive paleoseismic datasets,
such as those by \citet{benedetti_earthquake_2013} (or the correlation
may be negative as found by \citet{weldon_wrightwood_2004}), but the
number of paleoseismic datasets of sufficient size and quantity to
identify these effects with statistical significance are few indeed.

Because this modeling strategy involves sampling independence, it is
essentially a neutral model. If any correlation structure exists in the
sample sets, it will affect the displacement histories in predictable
ways. Negative autocorrelation in the sample sets, meaning that a long
interval (or slip distance) is followed by a short interval (or slip
distance) and vice versa, will cause a more rapid regression to the mean
slip rate line, and decrease the scatter in the slip rate estimates. A
positive correlation between recurrence (loading) intervals and slip
magnitudes will have the same effect. Conversely, positive
autocorrelation in either of the sample sets, or negative correlation
between the recurrence intervals and slip magnitudes, will lead to
slower regression to the mean line and therefore an increase in the
scatter of the slip rate estimates.

\subsection{Slip rate calculations}\label{slip-rate-calculations}

The uncertainty in the estimated slip rates due to earthquake cycle
variability is estimated by taking a function, \(\hat{R}\), that
calculates the mean slip rate within a time window \(t\), and sliding it
along the displacement series. \(\hat{R}\) is calculated simply as

\begin{equation}
  \hat{R}(D_0,D_1,t) = \frac{D_1 - D_0}{t}\; ,
  \label{eqn-rate-calc}
\end{equation}

where \(D_0\) is the cumulative displacement at the beginning of the
time window, \(D_1\) is the cumulative displacement at the end of the
time window, and \(t\) is the length of the time window. The \(\hat{}\)
symbol signifies an estimate rather than the true value \(R\). By
sliding \(\hat{R}\) over the displacement series, a set of many samples
of \(\hat{R}\) is generated, so that we may analyze the distribution.
The number of samples is \(n = N - t + 1\), where \(N\) is the length of
the total series (2,000,000 in this study).

A major goal of this study is to provide an answer to the question,
\emph{How long should slip rates be measured over in order to estimate a
meaningful rate?}\\
This question will be answered by looking at the distribution in
\(\hat{R}\) as a function of \(t\). Fifty values of \(t\) from 500 years
to 100,000 years, logarithmically spaced, are used. Note that given
\(\mu\) of 1000 years, this translates to 0.5-100 mean numbers of
earthquakes in the window.

The results of these calculations are shown in Figure
\ref{slip_rate_envelopes}. It is clear that the total variability in the
estimated slip rates is initially quite high when \(t\) is short
(\textless{}10,000 years or \textasciitilde{}10 earthquakes).
Particularly when \(t\) is \textless{}5,000 years, the maximum rates are
a factor of 3 or more greater than the true rate \(R\), but the median
rates are lower than \(R\)---this means it is more likely that fewer
earthquakes are captured in the time window than naively expected given
the mean recurrence. As the median is lower than \(R\), most
measurements over these short timescales will underestimate the mean
rate, although not necessarily by much.

With longer \(t\), between 10,000--20,000 years (or 10--20 earthquakes)
the variation in the slip rate estimates stabilize to within \(\pm\)
100\% of the mean (Figure \ref{slip_rate_envelopes}) for all
distributions, though this happens most quickly in the periodic
distribution, and most slowly in the clustered distribution. In fact,
the only exception here is that the lower bound of the clustered
distribution can stay at zero for more than 60 mean earthquake cycles.
For rate estimates longer than several tens of mean earthquake cycles,
the variation decreases very slowly but progressively with increasing
window length.

\subsubsection{Normalizing to different slip rates and earthquake
offsets}\label{normalizing-to-different-slip-rates-and-earthquake-offsets}

The distributions in this study were chosen to have \(\mu=1\) (kyr, m)
in order to make the mean slip rate \(R=1\) mm/yr, and therefore to make
all results easy to generalize to different systems with different real
rates. This normalization requires some values for the mean per-event
displacement \(\bar{D}\) and the slip rate \(R\), yielding a
normalization factor (or coefficient) that can be applied to the time
values shown in as the \(x\)-axis in Figure \ref{slip_rate_envelopes}:

\begin{equation}
  NF = \frac{\bar{D}}{R}
  \label{norm_eqn}
\end{equation}

\(NF\) is also equal to the mean recurrence interval \(\mu\) given
suitable unit transformations (though the recurrence interval may not be
known \emph{a priori}). For example, a fault with a slip rate of 5 mm/yr
but a per-event mean slip of 1 m has a normalization factor of 0.2,
meaning that earthquakes are 5 times as frequent on this fault as the
simulated fault, so the time window required for the rates to stabilize
is 0.2 times the simulated fault. For a fault with \(R=1\) mm/yr and
\(\bar{D}=2.5\) m, \(NF=2.5\), then the mean recurrence interval \(\mu\)
is 2.5 times as long as in these simulations, so \(NF=2.5\) and the
timescales for rate stabilization will be lengthened that much.

This normalization will obviously be more accurate if \(\bar{D}\) and
\(R\) are independently (and accurately) known or can be obtained from
other information. \(\bar{D}\) may be estimated paleoseismologically or
through the application of scaling relationships between fault length
and offset \citep{wells_new_1994,leonard_earthquake_2010}. The accuracy of \(\hat{R}\) is discussed
below, but suffice it for now to state that for more than
\textasciitilde{}10 earthquakes, \(\hat{R}\) should be acceptable.

\section{Discussion}\label{discussion}

\subsection{Interpreting measured
rates}\label{interpreting-measured-rates}

The most pragmatic motivation for this study is to understand how much
\emph{epistemic} uncertainty in a slip rate measurement results from the
\emph{aleatoric} uncertainty (or natural variability) in earthquake
recurrence. However, the previous results have focused on describing the
natural variability, and how much a measured rate may deviate from the
`true' secular rate, i.e. \(\hat{R}/R\). In these methods and results,
there is no epistemic uncertainty because all quantities are known
perfectly. Of course, in a real slip rate study, the measured value is
known, but the true value is not. The epistemic uncertainty then is
present, and can be quantified here by evaluating the true rate \(R\)
relative to the measured rates \(\hat{R}\), so that the distribution of
\(R / \hat{R}\) at a given \(t\) represents the epistemic uncertainty
distribution about the measured value.

The epistemic uncertainty relative to the measured rate is shown in
Figure \ref{epist_unc} for all distributions for the first 40,000 years
(or 40 earthquakes), represented by the 5, 25, median, 75 and 95
percentiles, and numerical results are given in Table
\ref{ep_unc_table}. Several things are clear in these plots:

First, the variance in the distributions is quite large for the first
several thousand years (or several mean earthquake cycles), but becomes
much more compact after \textasciitilde{}15 mean earthquake cycles, as
with the slip rate estimates in Figure \ref{slip_rate_envelopes}. The
right tails (or upper bounds in Figure \ref{epist_unc}) in fact are
infinite (or undefined) for the first few earthquake cycles, because in
some fraction of the simulations, \(\hat{R}\) is zero.

Second, the distributions are asymmetrical, especially the 5-95\%
interval. The 95th percentile is generally several times as far from the
measured value as the 5th percentile, meaning that the true value of the
slip rate may be much greater than the measured rate but not a
commensurately small fraction of the measured rate.

Third, the median rate before convergence at 5 \(\mu\) is greater than
the true rate, meaning that in most cases, very short-term slip rate
measurements will underestimate the true slip rate; this is a systematic
bias. This is a particular concern for paleoseismological slip rate
estimates, where there are rarely more than 5 events in any given
trench. However, the median is less than 2 times \(\hat{R}\) following
just 2-3 events, so the systematic bias is unlikely to be much greater
than the measurement uncertainties on the age or offset of the events.

\subsection{Evaluating slip rate
changes}\label{evaluating-slip-rate-changes}

It is of both theoretical and practical interest to be able to evaluate
whether fault slip rates may have changed over some time period, or
between multiple sets of measurements. From a theoretical perspective,
understanding under what conditions fault slip rates change can lead to
much insight into fault processes such as growth
\citep[e.g.]{roberts_fault_2002} and interaction
\citep[e.g.]{wallace_grouping_1987,dolan_long-range_2007}. Practically,
if an older (or longer-term) slip rate is quite different from the
contemporary rate, then its inclusion in a seismic hazard model may lead
to inaccurate hazard estimation.

First, a necessary definition: A slip rate change in this discussion
means a real change in \(R\), not a change in the estimate \(\hat{R}\),
which leads to a change in the distribution of earthquake recurrence
and/or displacement parameters with time (in statistical terminology,
the recurrence and displacement distributions are then
\emph{non-stationary}). This sort of change may be associated with
secular changes in fault loading, stemming from changing stress or
strain boundary conditions.

Discerning a real slip rate change, rather than a change in \(\hat{R}\)
due to natural variability, requires consideration of the lengths of
time over which the different slip rate measurements were made and the
associated uncertainty. If the distributions defined by two estimates
\(\hat{R_1}\) and \(\hat{R_2}\) and their empirical distributions
(reflecting the number of earthquakes as well as the \(COV\)s of the
underlying earthquake recurrence distributions are known, then the null
hypothesis that the two slip rates are drawn from the same stationary
distribution can be tested with a Kolmogorov-Smirnov test.

However, it is unlikely that the values for the recurrence distributions
and the number of earthquakes that have transpired are sufficently known
to make the calculation, unless the fault has received sufficient
paleoseismic and neotectonic study. As formal hypothesis testing may not
be possible given typical slip rate datasets, an informal way of gauging
the likelihood of a slip rate change is to crudely estimate
(`guesstimate') the number of possible earthquakes and the recurrence
distribution, and then use the closest values in Table
\ref{ep_unc_table} with propagated measurement uncertainty to evaluate
the amount of overlap between the two slip rate estimates. If the
overlap between the distributions is a small fraction of the total range
of the distributions, then it is likely that a real slip rate change
occurred. This is clearly not appropriate for a real hypothesis test,
but it may aid researchers in developing ideas or intuition about the
behavior of a given fault.

\conclusions

This work seeks to evaluate the effect of natural variability (aleatoric
uncertainty) in earthquake recurrence intervals on slip rate
measurements. The study simulates cumulative displacement during
2,000,000 earthquakes for a faults with stationary long-term slip rates
of 1 mm/yr and several different distributions for earthquake
recurrence, and then estimates the variation in estimated slip rates
over shorter time windows similar to those measured in
paleoseismological and neotectonic studies. The results display several
characteristics that are of importance to fault geologists seeking to
estimate slip rates on faults, or compare rates from different
measurement time windows or techniques:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The variability in slip rates calculated over time windows less than 5
  mean earthquake cycles is very large, but begins to stabilize
  following \textasciitilde{}10-20 earthquakes.
\item
  The most important factor in controlling the variability in slip rate
  estimates is the coefficient of variation (\(COV\)); the different
  distributions themselves are relatively unimportant. Faults with
  \emph{periodic} earthquakes, with \(COV < 1\), have less initial
  variability, and stabilize rapidly. Faults with \emph{unclustered}
  earthquakes, with \(COV = 1\), have more variability and stabilize
  more slowly. Faults with \emph{clustered} earthquakes,\\
  with \(COV > 1\), have a great amount of initial variability and
  require a large number of earthquakes to stabilize.
\item
  The epistemic uncertainties around a measured slip rate are similarly
  large initially and then decrease with time. These uncertainties are
  initially biased, such that the measured slip rates typically
  underestimate the true slip rate for \textless{}5 mean earthquake
  cycles, but this fades with time. However, the uncertainties remain
  asymmetric, with a strong right skew.
\end{enumerate}

\codeavailability{All code is available at 
https://github.com/cossatot/eq-slip-rate-variability-paper with an MIT 
license.}

\competinginterests{The author declares no competing interests.}

\bibliographystyle{copernicus}
\bibliography{eq_cycle_slip_rates}

\clearpage

\begin{figure}[t]
\includegraphics[width=8.3cm]{./figures/recurrence_dists.pdf}
\caption{Earthquake recurrence distributions. `logn' = lognormal. `exp' = 
  exponential. Colors for each distributions are the same in all figures. 
  \label{eq_rec_dists}}
\end{figure}

\clearpage

\begin{figure}[t]
  \includegraphics[width=8.3cm]{./figures/event_spacing.pdf}
  \caption{Spacing of 15 simulated successive earthquakes from each recurrence 
  distribution. Note that the gap between the last displayed earthquake and the 
  right side of the plot does not represent a long recurrence interval. 
  \label{eq_spacing}}
\end{figure}

\clearpage

\begin{figure}[t]
\includegraphics[width=8.3cm]{./figures/slip_dist.pdf}
\caption{Earthquake slip distribution. \label{slip_dist}}
\end{figure}

\clearpage

\begin{figure*}[t]
  \includegraphics[width=12cm]{./figures/disp_histories.pdf}
  \caption{Simulated displacement histories for each of the recurrence 
  distributions, and the `true' mean line at 1 mm/yr in black. \emph{A}: The 
  first hundred thousand years. \emph{B}: The entire 2 million years. The 
  histories are the same in both plots. \label{disp_histories}}
\end{figure*}

\clearpage

\begin{figure*}[t]
  \includegraphics[width=12cm]{./figures/slip_rate_envelopes.pdf}
  \caption{Envelopes of estimated slip rates as a function of the mean number 
  of earthquakes (or thousands of years) over which the slip rate was 
  estimated. All slip rates have a true value of 1 mm/yr. \emph{A}: periodic 
  distribution. \emph{B}: unclustered lognormal distribution. \emph{C}: 
  Clustered lognormal distribution. \emph{C}: Unclustered exponential 
  distribution. \label{slip_rate_envelopes}}
\end{figure*}

\clearpage

\begin{figure*}[t]
  \includegraphics[width=12cm]{./figures/epist_unc.pdf}
  \caption{Epistemic uncertainty relative to the measured rate for each of the 
  recurrence distributions, as a function of the mean number of earthquakes (or 
  thousands of years) over which the slip rates were measured. \emph{A}: 
  periodic distribution. \emph{B}: unclustered lognormal distribution. 
  \emph{C}: Clustered lognormal distribution. \emph{C}: Unclustered exponential 
  distribution.
  \label{epist_unc}}
\end{figure*}

\clearpage

\begin{table}[t]
\begin{tabular}{lllllll}
\tophline
distribution  &  $t$     &     5  &     25 &     50 &    75 &   95 \\
\middlehline
logn\_0.5 & 2531  &  0.51 &  0.79 &  1.13 &  1.77 &  4.37 \\
      & 4843  &  0.61 &  0.84 &  1.08 &  1.46 &  2.44 \\
      & 10,323 &   0.7 &  0.89 &  1.04 &  1.27 &  1.84 \\
      & 42,103 &  0.83 &  0.96 &  1.04 &  1.12 &  1.27 \\
logn\_1 & 2531  &  0.42 &  0.71 &  1.14 &  2.25 &   $\infty$ \\
      & 4843  &  0.51 &  0.75 &  1.07 &  1.67 &  5.62 \\
      & 10,323 &   0.6 &  0.82 &  1.04 &  1.38 &  2.55 \\
      & 42,103 &  0.72 &  0.89 &  1.03 &  1.18 &  1.55 \\
logn\_2 & 2531  &  0.35 &  0.68 &  1.35 &   $\infty$ &   $\infty$ \\
      & 4843  &  0.43 &   0.7 &  1.16 &  2.86 &   $\infty$ \\
      & 10,323 &  0.51 &  0.75 &  1.07 &  1.66 &   $\infty$ \\
      & 42,103 &   0.7 &  0.83 &  0.98 &  1.28 &  4.75 \\
exp\_1 & 2531  &   0.4 &   0.7 &  1.17 &  2.39 &   $\infty$ \\
      & 4843  &  0.49 &  0.75 &  1.08 &  1.67 &  4.79 \\
      & 10,323 &  0.61 &   0.8 &  1.03 &  1.37 &  2.32 \\
      & 42,103 &  0.76 &  0.89 &     1 &  1.15 &  1.43 \\
 \bottomhline
\end{tabular}
  \caption{Epistemic uncertainty table \label{ep_unc_table}}
\end{table}

\end{document}

### 2018-09-20 manuscript/compile_manuscript.sh

@@ -10,15 +10,18 @@ pandoc random_eq_slip_rate_variations.md --from markdown \
     --template=copernicus.template \
     --latex-engine=pdflatex \
     --filter pandoc-citeproc --bibliography=eq-cycle-slip-rate-paper.bib \
-    --csl=/Users/itchy/Zotero/styles/copernicus-publications.csl
+    --csl=/Users/itchy/Zotero/styles/copernicus-publications.csl \
+    --natbib \
+    --metadata link-citations=true
 
 
 elif [ "$OUT_TYPE" == "tex" ]; then
-pandoc random_eq_slip_rate_variations.md --from markdown \
+#pandoc random_eq_slip_rate_variations.md --from markdown \
+pandoc eq-slip-rate.md --from markdown \
     -o $OUTFILE_NAME$OUT_TYPE \
     --template=copernicus.template \
-    --filter pandoc-citeproc --bibliography=eq-cycle-slip-rate-paper.bib \
-    --csl=/Users/itchy/Zotero/styles/copernicus-publications.csl
+#    --filter pandoc-citeproc --bibliography=eq-cycle-slip-rate-paper.bib \
+#    --csl=/Users/itchy/Zotero/styles/copernicus-publications.csl
 
 elif [ "$OUT_TYPE" == "agu_tex" ]; then
 pandoc random_eq_slip_rate_variations.md --from markdown \

### 2018-09-20 scripts/slip_rate_fns.py

@@ -1,5 +1,6 @@
 import numpy as np
 import scipy.stats as ss
+import pandas as pd
 
 # sampling functions
 def lognormal_scale(arith_mean, arith_var):
@@ -11,16 +12,21 @@ def lognormal_shape(arith_mean, arith_var):
 
 
 def lognormal(mean, std, n_eqs=0, return_pdf=False):
-    shape = lognormal_shape(mean, std**2)
-    scale = lognormal_scale(mean, std**2)
-    
-    _lognorm = ss.lognorm(shape,
-                          scale=np.exp(scale))
-    
-    if return_pdf == True:
-        return _lognorm
+
+    if std == 0:
+        return np.ones(n_eqs) * mean
     else:
-        return _lognorm.rvs(n_eqs)
+
+        shape = lognormal_shape(mean, std**2)
+        scale = lognormal_scale(mean, std**2)
+        
+        _lognorm = ss.lognorm(shape,
+                              scale=np.exp(scale))
+        
+        if return_pdf == True:
+            return _lognorm
+        else:
+            return _lognorm.rvs(n_eqs)
 
 
 def exponential(mean, std, n_eqs=0, return_pdf=False):
@@ -39,11 +45,19 @@ def sample_eq_displacements(mean, std, n_eqs):
     return lognormal(mean, std, n_eqs)
 
 
+
+def sample_eq_displacements_bw(n_eqs):
+    bw_df = pd.read_csv('./bw_2006_offsets.csv')
+
+    return np.random.choice(bw_df.Dn.values, n_eqs)
+
+
 # This function makes a cumulative slip history given a recurrence mean,
 # standard deviation, recurrence interval distribution, and simulation
 # length.
 def make_cum_slip(rec_mean, rec_std, distribution, yrs=10000,
-                  displacement_mean=1, displacement_std=0.75):
+                  displacement_mean=1, displacement_std=0.75,
+                  use_biasi_weldon=False):
     n_eqs = int(yrs / rec_mean) + 100
     
     eq_rec_times = np.int_(distribution(rec_mean, rec_std, n_eqs))
@@ -52,9 +66,12 @@ def make_cum_slip(rec_mean, rec_std, distribution, yrs=10000,
     n_eqs = len(eq_dates)
     eq_rec_times = eq_rec_times[:n_eqs]
     
-    eq_disps = sample_eq_displacements(displacement_mean,
-                                       displacement_std,
-                                       n_eqs)
+    if use_biasi_weldon is False:
[…]

### 2018-09-20 manuscript/reviews/response_letter.md
added response letter [in progress]

# oskin

> [O1] This paper presents a useful thought experiment on the impact of 
> earthquake cycle variability on measured slip rates, and concludes that the 
> convergence on the expected value is a function of the coefficient of 
> variation. Overall this is a sensible conclusion. Underpinning this analysis 
> are four assumed variants of earthquake recurrence, and a function to express 
> the variability of slip per event.

> [O2] I would like to see the effect of COV isolated from the slip per event 
> distribution (use 1m slip for every event).

I performed this experiment; the figure with the slip rate results is shown
below for comparison with Figure 5 in the paper, and are included in a new
document in the supplemental material.

![Figure \label{fig}](../figures/nsv_slip_rate_envelopes.pdf)

The differences between the results of this experiment (fixed per-event
displacement of 1 m) and of the numerical experiment performed in the manuscript
are basically that these results are less smooth, but the total variance at any
point in time (*x-axis*) is less. This is simply due to removing the
stochasticity from one of the two variables in the system. The relative
spread in the data and convergence rates are unchanged. This is to be expected
as the variability in the per-event displacement is the same for all recurrence
distributions, so even though it is a random variable in the simulations, it is
not an experimental variable.

What I find the most interesting about this experiment is that the fluctuations
in the estimated slip rates show very clearly the mean earthquake cycles once
the noise from the per-event displacements has been removed. It is clear that
these are kind of damped or averaged out after ~7 earthquake cycles.
Nonetheless, though this is a cool pattern to see, I don't think it adds
enough insight to be worth including in the manuscript. If the reviewers and
editor think it is still of some value, we can put it in the supplemental
materials, but (as mentioned elsewhere) I prefer to keep the paper itself as
streamlined as possible.

**put table here.**

> [O3] I would also like to a see a more quantitative comparison of COV and a 
> convergence on the mean to back up the assertion that COV of the distribution 
> is more important than the distribution itself.

This comment is addressed in the response to [O18] below.

> [O4] The paper would be improved by a more quantitative, empirical basis and 
> discussion of physical processes that may drive such recurrence behavior.

This paper is written to be a quick and clean demonstration of the effects of 
recurrence (and slip) variability on slip rate estimations. I would prefer

> [O5] There is a literature of ideas to draw upon, such as post-seismic fault 
> reloading (Kenner and Simons, 2005), earthquake super cycles (Sieh et al., 
> 2008; Weldon et al., 2004), isolated versus fault-network behavior (Berryman et 
> al., 2012). Some of these ideas are discussed briefly but need more 
> explanation.

I don't think that a more in-depth discussion of these ideas is beneficial to 
the goals of the paper. As stated in the response to [04], I am simply 
demonstrating the effects of variability in the earthquake process to slip rate 
studies.

It is clear that this variability exists. But it is not clear

> [O6a] Likewise one could examine actual earthquake slip distributions (not 
> landform offsets of historic events, which convolve landscape processes with 
> tectonic slip) to develop an empirical basis for the slip function.

Such a distribution is given by Biasi and Weldon (2009), following work done by 
Hemphill-Haley and Weldon (1999). It is a bit different than the lognormal 
distribution used in that the probability of relatively low values (zero or 
near-zero) is higher than in a lognormal distribution. The sample COV of this 
distribution is 0.67, slightly lower than the lognormal slip distribution used
in the paper, with a COV of 0.75.

As an experiment, I have re-done the simulation sampling from this 
distribution; the data are given as 1313 discrete points from earthquakes 
worldwide, normalized to the mean slip per event. I have sampled randomly from 
this finite set, with replacement, instead of interpolating the set into a 
continuous distribution and sampling from that. The results are in the new
supplemental materials.

The results are very similar to the results in the original manuscript. The 
5-95% interval is slightly narrower than

These

> [O6b] Some of the scatter in slip distributions is likely due to 
> underreported measurement uncertainty (Gold et al., 2013), and thus the 
> cancellation of this error over multiple earthquakes should let cumulative 
> slip converge more quickly than may be predicted from the author’s model.

Either I don't understand this comment (which is quite possible) or it is a bit 
misapplied. The only reading of this comment under which one expects faster 
convergence than I have modeled is if the reviewer believes that measurement 
error is some component of the total variability represented. But that is not 
the case in the modeling; these distributions are taken to represent only 
aleatory variability and the study is performed with assumptions of zero 
measurement error.

> [O7] Page 1, line 4. The open interval problem is well known and attempts to 
> quantify it do exist on case-by-case basis.

Yes, and some of these cases are cited in the introduction. However, the 
open-interval problem simply deals with the uncertainty in a single recurrence 
interval (the present one), and not the variability that is present throughout 
all of the closed earthquake intervals that have contributed to the measured 
offset; this larger issue is the topic of the manuscript.

*No changes.*

> [O8] Page 1, line 13. It seems odd to characterize uncertainty due to a 
> random distribution as epistemic. Isn’t this unreported aleatory uncertainty?

No, it's not unreported aleatory uncertainty. It is aleatory variability 
*transformed* into epistemic uncertainty by a condition of ignorance of the 
underlying distribution.

The issue is that, with a single measurement (such as a slip rate measurement) 
one doesn't know what the mean slip rate is, or how much this rate varies 
through time.

A quick dice analogy:

If I have a fair, six-sided die, and I

> [O9] Page 1, line 20: Why is marker in quotes?

> [O10] Page 2, line 5. afterslip and creep also contribute.

> [O11] Page 2, line 10-11. Awkward sentence. Break into two.

> [O12] Page 2, line 13 and other citations:  persistent use of ‘e.g’ after 
> citing only one or two articles is poor form and makes this reader think that 
> the author has not adequately explored the literature.

The use of 'e.g.' denotes that the given citations are not authoritative in the 
sense that the cited works are where the concepts given are first introduced or 
best developed, as this isn't true. The cited works are generally just modern, 
high quality studies that exemplify the topic at hand.

I don't really care what readers may think of the depth of my scholarship.

> [O13a] Page 3,  line 11.  This is not the correct definition of an 
> exponential / poisson distribution.  There is no prescribed number of events, 
> only a prescribed time-independent probability.

The time-independent probability is the mean rate of events. The mean rate of 
events is the mean number of events that occur within some time interval.

Obviously with finite sample sets (of time, or of events) there will be some 
variation--otherwise I wouldn't have written the paper.

> [O13b] It is also worth noting that this is physically unrealistic at short 
> time intervals because it violates elastic rebound.

Elastic rebound is a theory, not a law. It is unfortunately a step removed from 
modern understanding of the mechanics of earthquakes, which are based around 
stress, not strain. These map to each other nicely in the case of elastic and 
Newtonian viscoelastic rheologies, but not as nicely with rate-dependent 
rheologies, which are often considered the best characterization of the lower 
crust and upper mantle (e.g., Hetland and Hager, xxx). It's also hard to put 
strain in a framework with friction, for example. 'Physically realistic' 
modeling has to make a lot of assumptions and use heavy duty equipment (finite 
elements, for example) to incorporate strain.

There is also a separate issue with elastic rebound theory: It's not very easy 
to tell whether all the accumulated shear strain was released in an earthquake 
or not. What kind of measurements would tell us this?

I strongly suspect that we are fundamentally underestimating the frequency of 
very short recurrence intervals on faults. They're close to invisible to 
paleoseismology, which is our main source of data for recurrence interval 
statistics, because very closely-spaced events may not each produce 
differentiable colluvial wedges or other signs of surface deformation. This 
could plausibly result in a strong sample bias in the statistics. Nonetheless, 
we have clear observations of short recurrence intervals in the past few years. 
For example, some parts of the Monte Vettore fault slipped about 20 cm in the 
Amatrice earthquake and then ~2 m in the Norcia earthquake a few months later 
(Gruppo di Lavoro INGV sul Terremoto di Amatrice-Visso. (2016, October 29). 
PRIMO RAPPORTO DI SINTESI SUL TERREMOTO DI VISSO ML 5.9 DEL 26 OTTOBRE 2016 
(ITALIA CENTRALE). Zenodo. http://doi.org/10.5281/zenodo.163818).

From a fault mechanics perspective, some researchers (for example Mark Zoback 
and his students, primarily Townend, as well as myself) believe that the shear 
stress on a fault required to initiate failure is much greater than the stress 
drop during the event, i.e. shear stress does not go to zero. Failure is 
decently described by Mohr-Coulomb models, and at, say, 10 km depth, the 
confining pressure is almost 300 MPa. With reasonable static friction 
coefficients (say 0.3 to 0.6) and typical pore fluid pressure,

> [O14] Page 3, line 24. It would be useful to briefly discuss how shape and 
> scale affect distributions generally. Shape governs the how tailed and is 
> dimensionless; scale determines the spread of the distribution and is 
> dimensioned (in years for this case).

Isn't this implicit in the

> [O15] Page 4, line 6.  Pareto distribution is another, simpler distribution 
> needing only shape and scale to describe COV > 1

The Pareto distribution isn't used to describe earthquake recurrence, as far as 
I know. I believe that its only use in seismology is the tapered Pareto 
distribution for magnitude-frequency distributions by Yan Kagan (and perhaps 
others); this is similar in concept but not close enough to use here.

(Moderately) interestingly, I gather that Pareto/power-law distributions often 
form in environments characterized by exponential growth and aggregation 
(population, money, etc.), while the somewhat similar and discussed 
hyperexponential distribution can result from the sum of possibly-independent 
random Poisson processes, though both give rise to final distributions with 
similar tails.

*No changes.*

> [O16] Page 4, line 15. Akciz et al (2012) revised Grant and Sieh (1994) and 
> found much more periodic behavior.

The referenced sentence simply states "However, one can find examples of 
studies indicating the opposite conclusions", to reinforce the paragraph's 
opening statement that "No consensus exists among earthquake scientists as to 
the most appropriate recurrence interval distribution."  That a study revised a 
previous study and found different results further reinforces this point, but I 
don't think there is additional benefit to citing it.

*No changes.*

> [O17] Page 4, line 28.  The author should consider non-dimensionalizing the 
> results of this study to facilitate more general use of its results. Instead 
> of mean slip of 1m, one would refer to non-dimensional slip of 1 and multiply 
> by average slip per event to scale the results.  This is effectively what the 
> author describes already, though without formal non-dimensionalization.

I considered formal nondimensionalization but decided against it.

In the end, I decided to present the results dimensionally because most 
geologists think dimensionally (myself included).

Dimensional thinking allows for different heuristics to be used when analyzing 
methods and results, than non-dimensional thinking. Basically, 
non-dimensionalizing parameters forces parameters to only be defined in terms 
of their relationship with each other, and to only exist within the context of 
this specific problem.

If I say a fault has a mean slip of 1, and a mean recurrence interval of 1, and 
a mean slip rate of 1, it's hard to picture such a fault. Are those reasonable 
values for these parameters, individually? Can't say. Furthermore, the slip 
rate isn't even 1, it's 1/1, because it's a rate and the non-dimensional units 
are still not arithmetically compatible units--multiplication and division are 
possible (kind of, but not reducible) but addition and subtraction are not at 
all defined. (Please forgive me for not knowing enough measure theory to 
rigorously state this...).

Non-dimensionalization in many geological problems reduces the clarity of the 
analysis or solution because it strips it of context. At the same time it can 
facilitate the manipulation of the equations within the study, or during 
programming, etc. Physicists like it because it makes their work easier, I 
would say, and they're all pretty used to it.

Compare this to the dimensionalized problem in the paper. Can you imagine a 
fault with a mean per-event slip of 1 m, a recurrence interval of 1000 years, 
and a slip rate of 1 mm/yr?  I can, and I can place it--it's a small but pretty 
active intraplate fault, or a splay in a plate boundary. It's the kind of fault 
one might actually go and trench.

However, all the numbers are 1 (or 1000) so that it is easy to scale to other 
faults by multiplying by non-dimensional scaling factors (i.e., for a Mongolian 
fault with a slip rate of 3 mm/yr and a per-event slip of 10 m, you scale the 
slip by 10 and the slip rate by 3). It's not accidental.

*No changes.*

> [O18] Page 5, line 15. The statement ‘appears to be related to COV’ is 
> disappointing. Given that this paper is entirely simulation, the author 
> should be able to make a quantitative comparison of slip-rate variance to 
> COV.

Upon reflection this basically comes out of the mathematical definition of 
standard error:

> [O19] Page 5, line 29. This problem has been studied (Weldon et al., 2004; 
> Sieh et al., 2008). The Sumatran subduction zone work is particularly 
> relevant and completely overlooked here.

The work by Weldon et al. (2004) is referenced on Page 6, Line 3.
Neither of these papers deal with the of autocorrelated recurrence intervals in 
any quantitative or otherwise explicit manner. One can read both papers and not 
get a sense of whether a short recurrence interval implies the next recurrence 
interval will be short or long. Both papers deal with the concept of 
'earthquake supercycles' but these are not defined in any quantitative sense.

[Note: I've analyzed Weldon's data (from K. Scharer's refined OxCal earthquake 
dates) and found that earthquake recurrence interval duration at the Wrightwood 
and Pallett Creek sites has a negative autocorrelation, i.e. a short recurrence 
interval is likely to be followed by a long recurrence interval, and vice 
versa. This is quite unlike the autocorrelation in ,

> [O20] Page 6, line 1. Zero friction at rupture arrest is very unrealistic, 
> and not a prerequisite for characteristic behavior.

It is not a necessary condition but it is a sufficient one, when coupled with 
fairly regular reloading and failure conditions. I personally think it's 
unlikely as well, but zero (or very near zero) friction is part and parcel with 
complete stress drop in major earthquakes, which is supported by many studies 
[REFS]. Furthermore, there are a host of laboratory experiments which suggest 
that friction during slip decreases to very low values [MORE REFS].

### 2018-09-20 manuscript/eq_cycle_slip_rates.tex
new sims in revisions

@@
 There are a handful of statistical models for earthquake recurrence
-interval distributions that are in widespread consideration.
+interval distributions that are under widespread consideration by the
+seismological community.
 
@@
 associated with a Poisson process, and is the distribution that results
-from a prescribed number of earthquakes being distributed uniformly
-randomly within some time interval. Consequently, the probability of an
-earthquake (or other event) occurring at any time does not change with
-time since the previous event (in other words, the hazard function is
-time-invariant); this leads to the characterization of the exponential
-recurrence distribution as `random', `memoryless' or `time-independent'.
-The exponential distribution is also the simplest to describe
-statistically, as it requires only one parameter (the mean rate
-parameter), which is a statistical scale parameter. The standard
-deviation of a large number of samples generated from an exponential
-distribution is equal to the mean.
+from earthquakes being distributed uniformly randomly within some time interval.
+Consequently, the probability of an earthquake (or other event) occurring at any
+time does not change with time since the previous event (in other words, the
+hazard function is time-invariant); this leads to the characterization of the
+exponential recurrence distribution as `random', `memoryless' or
+`time-independent'. The exponential distribution is also the simplest to
+describe statistically, as it requires only one parameter (the mean rate
+parameter), which is a statistical scale parameter. The standard deviation of a
+large number of samples generated from an exponential distribution is equal to
+the mean.
 
@@
 some variability. This is representative of behavior observed in many
-studies \citep[e.g.]{zielke_slip_2010,klinger_characteristic_2011,zielke_earthquake_2018}. Taken together, the mean slip of 1 m and the
+studies \citep[e.g.]{zielke_slip_2010, klinger_characteristic_2011,
+  zielke_earthquake_2018}. Taken together, the mean slip of 1 m and the
 mean recurrence interval of 1000 years shared by each of the recurrence
@@
 be generalized to faults with different parameters.
+
+The choice of the lognormal distribution is for convenience, simplicity, and
+flexibility: it is a common, well-known distribution and--should one be
+interested--can be easily given different shape and scale values to modify the
+\(COV\) or change the mean slip rate in the modeling code used in this paper.
+However, it is not necessarily the most accurate representation of earthquake
+slip variability. **BIASI AND WELDON (2006)** compiled field measurements of
+surface ruptures from 13 earthquakes, and the resulting distribution has some
+significant differences with the lognormal distribution used here, though the
+\(COV\) of 0.67 is quite close to the value used here. To test the
+sensitivity of the results given in this paper to the choice of slip
+distribution, the numerical simulations in presented in this work were run with
+the only change being the use of the empirical slip distribution from **BIASI
+AND WELDON (2006)**, and the results are given in the supplemental material.
+Though there is more discussion there, the results are essentially identical to
+those presented below.
 

### 2018-10-14 manuscript/reviews/response_letter.md
updated all mods

@@
-# oskin
+ oskin
 
@@
 below for comparison with Figure 5 in the paper, and are included in a new
-document in the supplemental material.
+document in the supplemental material. The most relevant figure is also shown
+here:
 
@@
 Nonetheless, though this is a cool pattern to see, I don't think it adds
-enough insight to be worth including in the manuscript. If the reviewers and
-editor think it is still of some value, we can put it in the supplemental
-materials, but (as mentioned elsewhere) I prefer to keep the paper itself as
-streamlined as possible.
-
-**put table here.**
+enough insight to be worth including in the manuscript. I have added the figure
+and table that show the slip rate variation through time to a new supplemental
+document, accompanied by a brief discussion.
 
@@
 This paper is written to be a quick and clean demonstration of the effects of 
-recurrence (and slip) variability on slip rate estimations. I would prefer
+recurrence (and slip) variability on slip rate estimations. I want to keep the
+work as simple and as digestible as possible, for a few reasons:
+
+1. The intended audience for the paper is not only crustal deformation
+   researchers, but others in the seismic hazard community as well--this
+   includes engineers, geotechnical workers, analysts in the insurance industry,
+   etc. In my experience as a member of this community, many others are
+   only interested in these sorts of phenomena to the degree that they are
+   consequential and actionable; their intellectual interests are often oriented
+   towards their fields of expertise (structural engineering, ground motions,
+   human and economic exposure, etc.). I want this paper to be a straightforward
+   reference for how to evaluate slip rate data in light of aleatory variability
+   that is not tied down in jargon or linked to specific geological or
+   geophysical models or ideas that may not stand the test of time. Because of
+   how variable and poorly-understood earthquake recurrence and fault
+   interaction phenomena are, an in-depth discussion without resolution may well
+   be off-putting to much of the audience that I would like to read this paper.
+
+2. I don't believe that the state of the science is such that several paragraphs
+   to a page of discussion can summarize our (minimal) understanding of these
+   issues with any degree of satisfaction. Understanding earthquake recurrence
+   and fault interaction is a fascinating research topic, and it deserves either
+   a very full treatment (a review paper, lots of data analysis and modeling) or
+   simply a few references before the paper moves on. We are not at the stage
+   where we have a lot of answers, so it is very easy to get bogged down.
 
@@
 
-I don't think that a more in-depth discussion of these ideas is beneficial to 
-the goals of the paper. As stated in the response to [04], I am simply 
-demonstrating the effects of variability in the earthquake process to slip rate 
-studies.
-
-It is clear that this variability exists. But it is not clear
+I disagree. It is clear that this stochasticity exists in nature, but I don't
+think that we understand why. The modeling by Brad Hager's and Mark Simons'
+research groups
 
@@
 continuous distribution and sampling from that. The results are in the new
-supplemental materials.
+supplemental materials (Figures S1 and S2, Table S1, and some discussion).
 
-The results are very similar to the results in the original manuscript. The 
-5-95% interval is slightly narrower than
-
-These
+The results are nearly indistinguishable.
 
@@
 
-Upon reflection this basically comes out of the mathematical definition of 
-standard error:
-
 > [O19] Page 5, line 29. This problem has been studied (Weldon et al., 2004; 
@@
 
-The work by Weldon et al. (2004) is referenced on Page 6, Line 3.
-Neither of these papers deal with the of autocorrelated recurrence intervals in 
-any quantitative or otherwise explicit manner. One can read both papers and not 
+The work by Weldon et al. (2004) is referenced on Page 6, Line 3. Neither of
+these papers deal with the topic of autocorrelated recurrence intervals in any
+quantitative or otherwise explicit manner, and I don't read anything that I can
+interpret as a qualitative discussion either. One can read both papers and not
 get a sense of whether a short recurrence interval implies the next recurrence
-interval will be short or long. Both papers deal with the concept of 
-'earthquake supercycles' but these are not defined in any quantitative sense.
+interval will be short or long, much less any quantification. Both papers deal
+with the concept of 'earthquake supercycles' or groups of earthquakes that are
+relatively tightly-spaced and separated from other groups by long recurrence
+intervals. This may share a conceptual link, but from a technical perspective
+this likely has more to do with periodicity than autocorrelation, and these are
+mathematically separate such that a periodic sequence may have positive or
+negative autocorrelation. Goh and Barabasi (2008, Europhysics Letters) is a
+useful discussion on the topic of autocorrelation vs. periodicity in regards to
+quantifying clustering behavior (though one may safely ignore the references to
+seismicity in that paper).
 
@@
 interval is likely to be followed by a long recurrence interval, and vice 
-versa. This is quite unlike the autocorrelation in ,
+versa. This is quite unlike the autocorrelation in the Puget Lowlands of
+Washington State, which is more like a network of generally similar faults
+instead of an isolated or dominant plate boundary fault like the SAF or the
+Alpine Fault as referenced in another comment on Kelvin Berryman's work, where a
+long recurrence interval will more likely be followed by another long recurrence
+interval. I don't know what this means, or whether it's all noise, but it's
+intriguing to me and I put this bit in the paper in hopes of catching the
+attention of others who may be looking for a project.]
 
@@
 fairly regular reloading and failure conditions. I personally think it's 
-unlikely as well, but zero (or very near zero) friction is part and parcel with 
-complete stress drop in major earthquakes, which is supported by many studies 
-[REFS]. Furthermore, there are a host of laboratory experiments which suggest 
-that friction during slip decreases to very low values [MORE REFS].
+unlikely as well (see Styron and Hetland 2015 for example), but zero (or very
+near zero) friction is part and parcel with complete stress drop in major
+earthquakes, which is supported by many studies [REFS]. Furthermore, there are a
+host of laboratory experiments which suggest that friction during slip decreases
+to very low values [MORE REFS].

### 2018-10-14 manuscript/eq_cycle_slip_rates.tex
updated all mods

@@
 Fault slip rates are generally estimated by dividing measurements of the
-offset of geologic `marker' features by the time over which that offset
+offset of geologic marker features by the time over which that offset
 accumulated. The uncertainty in the resulting slip estimate is typically
@@
 particularly given successive events from the tails of the recurrence
-interval or displacement distributions. Though the physical mechanisms
+interval or displacement distributions.
+
+The physical mechanisms responsible for the aleatoric variability in earthquake
+recurrence intervals and displacements are still unclear, and the subject of
+active investigation. Most earthquakes serve to release differential stresses
+caused by relative motions of tectonic plates or smaller crustal blocks;
+relative plate velocities measured over tens of thousands to millions of years
+from geologic reconstructions are similar enough to those measured over a few
+years through GPS geodesy that sudden, transient accelerations and decelerations
+are unlikely (e.g., **DEMETS AND DIXON 1999**). As a consequence, some plate
+boundary faults are among the most regularly-rupturing faults known,
+particularly sections that are isolated from nearby faults
+\citep[e.g.]{berryman_major_2012} and therefore not affected by stress
+perturbations resulting from earthquakes on other faults. These stress changes
+may be 'static' coseismic instantaneous stresses in the elastic upper crust resulting from
+earthquake displacement, or analogous stress changes in the viscoelastic lower
+crust or upper mantle from the time-dependent relaxation of static stress
+perturbations, or 'dynamic' transient stress changes that accompany the passage
+of seismic waves from nearby or distant earthquakes. Regardless of the nature
+
+Intraplate faults
+
+It is important to keep in mind that each earthquake in a sequence places
+points of rock on both sides of the fault that were not previously in contact.
+
+Though the physical mechanisms
 responsible and the statistical character of this natural variability
@@
 erosion, nor does it consider any measurement uncertainty in the age or
-offset of the faulted geologic markers; these measurements are assumed
-to be perfect. Additionally, though natural variability in
+offset of the faulted geologic markers; these quantities are assumed
+to be perfectly known. Additionally, though natural variability in
 per-earthquake displacement is included in the model, it is minor and
-the same for all recurrence distributions. Furthermore, though the model
+the same for all recurrence distributions; though it is a random variable in the
+simulations, it is not an experimental variable. Furthermore, though the model
 has one length dimension (fault offset), it is still best thought of as
@@
 \(COV\) or change the mean slip rate in the modeling code used in this paper.
+
 However, it is not necessarily the most accurate representation of earthquake
 slip variability. **BIASI AND WELDON (2006)** compiled field measurements of
-surface ruptures from 13 earthquakes, and the resulting distribution has some
-significant differences with the lognormal distribution used here, though the
-\(COV\) of 0.67 is quite close to the value used here. To test the
+surface ruptures from 13 earthquakes. The resulting distribution (Figure S1) has
+some significant differences with the lognormal distribution used here, though
+the \(COV\) of 0.67 is quite close to the value used here. To test the
 sensitivity of the results given in this paper to the choice of slip
@@
 the only change being the use of the empirical slip distribution from **BIASI
-AND WELDON (2006)**, and the results are given in the supplemental material.
-Though there is more discussion there, the results are essentially identical to
-those presented below.
+AND WELDON (2006)**, and the results are given in the supplemental material
+(Figure S2, Table S1). Though there is more discussion in the supplements, the
+results are essentially identical to those presented below. As an additional
+experiment, the numerical simulations have been run using an invariant per-event
+displacement of 1 m. Though this is not a realistic scenario, it allows for a
+deconvolution of the effects of earthquake time stochasticity and earthquake
+displacement stochasticity. The results are shown in Figure S3 and Table S2, and
+discussed in the supplemental materials; there are noticeable differences in the
+results, but they are small and do not call into question the results and
+conclusions presented below.
 
@@
 
-This normalization will obviously be more accurate if \(\bar{D}\) and
-\(R\) are independently (and accurately) known or can be obtained from
-other information. \(\bar{D}\) may be estimated paleoseismologically or
-through the application of scaling relationships between fault length
-and offset \citep{wells_new_1994,leonard_earthquake_2010}. The accuracy of \(\hat{R}\) is discussed
-below, but suffice it for now to state that for more than
+This normalization will obviously be more accurate if \(\bar{D}\) and \(R\) are
+independently (and accurately) known or can be obtained from other information.
+\(\bar{D}\) may be estimated paleoseismologically or through the application of
+scaling relationships between fault length and offset
+\citep{wells_new_1994,leonard_earthquake_2010}. The accuracy of \(\hat{R}\) is
+discussed below, but suffice it for now to state that for more than
 \textasciitilde{}10 earthquakes, \(\hat{R}\) should be acceptable.

### 2018-10-14 manuscript/reviews/response_letter.md
finished revisions

@@
- oskin
+---
+title: Response to Reviewers, "The impact of earthquake cycle variability on 
+neotectonic and paleoseismic slip rate estimates"
+author:
+  name: Richard Styron
+---
+
+## Review by Michael Oskin
 
@@
 
+*Changes: New experiment added to new supplemental materials, and brief 
+discussion in the main manuscript.*
+
 > [O3] I would also like to a see a more quantitative comparison of COV and a 
@@
 
-This paper is written to be a quick and clean demonstration of the effects of 
-recurrence (and slip) variability on slip rate estimations. I want to keep the
-work as simple and as digestible as possible, for a few reasons:
+I don't understand how the work could be more quantitative--it's a purely 
+numerical study.
+
+As for making it more empirical, there are several ways to do this:
+
+1. *Using empirical distributions for earthquake recurrence and slip.* There 
+   isn't a lot of consensus on empirical (non-parametric) earthquake recurrence 
+   distributions; most of the community prefers parametric distributions such 
+   as the lognormal, Weibull or Brownian Passage Time distributions. As 
+   discussed in the paper by Matthews et al. (2002) that is the most prominent 
+   introduction to the Brownian Passage Time distribution, given the very small 
+   number of samples for earthquake recurrence that we will have for a given 
+   section of faulting, it is impossible to discriminate between these 
+   distributions, so going with the lognormal (as I have done in this study) is 
+   justifiable on empirical grounds as well as practical ones (it is familiar, 
+   implemented in many programming environments, and easy to manipulate). With 
+   regards to empirical slip rate distributions, I have added an experiment 
+   that uses one, which is explained in more detail in response to comment 
+   [O6a] below.
+
+2. *Going through the literature and evaluating studies that claim that slip 
+   rates have changed (or have not changed) in light of the work presented 
+   here.* I considered this, and in fact a major motivation for me to begin the 
+   analysis was my skepticism over some recent literature claiming major slip 
+   rate changes over relatively short timespans. However, I opted not to do 
+   this in the paper, mainly because I didn't want to pick fights. There is a 
+   bit of a paradox here: If I claim that the conclusions in a paper claiming 
+   secular slip rate changes (or fluctuations) are actually due to aleatoric 
+   variability in earthquake recurrence, I will probably anger those authors 
+   and decrease the likelihood that they will consider these results in 
+   subsequent work. I'd rather write a more toothless paper that doesn't single 
+   out any given researchers, and is therefore a bit easier to swallow by all.
+
+3. *Incorporating measurement uncertainty.* Measurement uncertainty is a very 
+   large factor that affects the results of any slip rate measurements, and I 
+   fully agree with comment [O6b] that it is in most cases underreported, both 
+   in offset measurements (as that comment references) and in geochronologic 
+   dating of any sort.  I chose to leave it out of this paper because I really 
+   wanted to focus on the aleatoric variability, which is generally 
+   *underappreciated* as opposed to *underreported*.
+
+Per a discussion of the physics and mechanics behind recurrence behavior: I 
+have added a short discussion, but I don't want to really dig into the topic, 
+for two reasons:
 
@@
 
-2. I don't believe that the state of the science is such that several paragraphs
-   to a page of discussion can summarize our (minimal) understanding of these
-   issues with any degree of satisfaction. Understanding earthquake recurrence
-   and fault interaction is a fascinating research topic, and it deserves either
-   a very full treatment (a review paper, lots of data analysis and modeling) or
-   simply a few references before the paper moves on. We are not at the stage
-   where we have a lot of answers, so it is very easy to get bogged down.
+2. I don't think that we have a great understanding of the real mechanisms, 
+   yet. There are a variety of mechanisms under consideration (e.g., co- and 
+   post-seismic elastic and viscoelastic Coulomb stress changes, stress 
+   transients, dynamic triggering, pore fluid pressure fluctuations, 
+   fluctuations in the frictional failure threshold on a fault) in addition to 
+   actual secular changes in tectonic loading rates. The time-dependent 
+   mechanisms (particularly post-seismic processes) often show different 
+   behavior with regard to whether they are 'spun-up' and at a dynamic 
+   equilibrium, or not. And all of these mechanisms are necessarily linked to 
+   uncertainty as to how (and where) faults are loaded to begin with--whether 
+   the loading is in the elastic crust, in the viscoelastic/viscous mid or 
+   lower crust (in a continuum style), or on a discrete creeping dislocation 
+   down-dip of the brittle fault. There is a big range of scientific opinion on 
+   all of these questions. As a community we are begging for a big review paper 
+   to at least concatenate and organize these ideas and potentially test them 
+   or at least sort them into compatible vs. mutually exclusive sets for future 
+   testing. But we don't have that right now, so the topic is kind of a big 
+   mudhole. I will dip my toe in but I really want to avoid falling in for the 
+   purposes of this paper.
+
+*Changes: New experiment with empirical slip distribution, and discussion of 
+physical mechanisms behind aleatory recurrence variability.*
 
@@
 > reloading (Kenner and Simons, 2005), earthquake super cycles (Sieh et al., 
-> 2008; Weldon et al., 2004), isolated versus fault-network behavior (Berryman et 
-> al., 2012). Some of these ideas are discussed briefly but need more 
+> 2008; Weldon et al., 2004), isolated versus fault-network behavior (Berryman 
+> et al., 2012). Some of these ideas are discussed briefly but need more 
 > explanation.
 
-I disagree. It is clear that this stochasticity exists in nature, but I don't
-think that we understand why. The modeling by Brad Hager's and Mark Simons'
-research groups
+I have added a short discussion (two paragraphs) on the topic, but as noted in 
+my response to [O4] I don't think that a more full discussion is warranted. I 
+want this paper to be a simple, easy-to-digest paper and I think that a long 
+and necessarily unsatisfying (as we don't have answers yet) discussion on the 
+mechanisms behind recurrence variability will be an obstacle, and many readers 
+will just put the paper down.
+
+*Changes: discussion added.*
 
@@
 
+*Changes: new experiment added.*
+
 > [O6b] Some of the scatter in slip distributions is likely due to 
@@
 
+*Changes: None.*
+
 > [O7] Page 1, line 4. The open interval problem is well known and attempts to 
@@
 
-*No changes.*
+*Changes: None..*
 
@@
 
-No, it's not unreported aleatory uncertainty. It is aleatory variability 
-*transformed* into epistemic uncertainty by a condition of ignorance of the 
-underlying distribution.
+Aleatory and epistemic uncertainty are not mutually exclusive categories. Much 
+epistemic uncertainty results from aleatoric variability, particularly when the 
+underlying distributions that characterize the aleatoric variability are not 
+known.
 
-The issue is that, with a single measurement (such as a slip rate measurement) 
-one doesn't know what the mean slip rate is, or how much this rate varies 
-through time.
+This is one of those instances: The framing of the situation is that one has 
+made a single slip-rate 'measurement' (net slip / time) without knowledge of 
+where the fault is in its earthquake cycle, what the past earthquake history 
+is, and what the distributions of slip and recurrence are for that fault to 
+begin with. Thus the condition is one of ignorance, i.e. epistemic uncertainty, 
+and this section of the study shows how to approximately quantify this 
+uncertainty under different assumptions of the slip and recurrence 
+distributions.
 
-A quick dice analogy:
-
-If I have a fair, six-sided die, and I
+*Changes: None.*
 
@@
 
+I wanted to declare that it was a technical term and not a word that I 
+arbitrarily applied to the situation. But this isn't necessary.
+
+*Changes: Quotes removed.*
+
 > [O10] Page 2, line 5. afterslip and creep also contribute.
 
+Truth.
+
+*Changes: Afterslip and creep added to sentence.*
+
 > [O11] Page 2, line 10-11. Awkward sentence. Break into two.
+
+Ok.
+
+*Changes: Sentence broken.*
 
@@
 
-The use of 'e.g.' denotes that the given citations are not authoritative in the 
-sense that the cited works are where the concepts given are first introduced or 
-best developed, as this isn't true. The cited works are generally just modern, 
-high quality studies that exemplify the topic at hand.
+The use of 'e.g.' denotes that the given citations are not authoritative or 
+canonical in the sense that the cited works are where the concepts given are 
+first introduced or best developed, as this isn't true. The cited works are 
+generally just modern, high quality studies that exemplify the topic at hand.
 
 I don't really care what readers may think of the depth of my scholarship.
+
+For what it's worth, 'e.g.' should be before the references but was placed 
+after by a LaTeX bug that I hadn't diagnosed.
 
@@
 
+Nevertheless, the statement actually made is that the spacing between uniform 
+random samples in some interval is characterized by an exponential 
+distribution, which is true. It is not stated that this is the definition of 
+the distribution.
+
+*Changes: None.*
+
 > [O13b] It is also worth noting that this is physically unrealistic at short 
@@
 
-Elastic rebound is a theory, not a law. It is unfortunately a step removed from 
-modern understanding of the mechanics of earthquakes, which are based around 
-stress, not strain. These map to each other nicely in the case of elastic and 
-Newtonian viscoelastic rheologies, but not as nicely with rate-dependent 
-rheologies, which are often considered the best characterization of the lower 
-crust and upper mantle (e.g., Hetland and Hager, xxx). It's also hard to put 
-strain in a framework with friction, for example. 'Physically realistic' 
-modeling has to make a lot of assumptions and use heavy duty equipment (finite 
-elements, for example) to incorporate strain.
+Elastic rebound is a hypothesis, not a law, and is phenomenological instead of 
+physical in nature. It is unfortunately a step removed from the modern 
+understanding of the mechanics of earthquakes, which are based around stress, 
+not strain. These map to each other nicely in the case of elastic and Newtonian 
+viscoelastic rheologies, but not as nicely with rate-dependent rheologies, 
+which are often considered the best characterization of the lower crust and 
+upper mantle (e.g., Hetland and Hager, 2006). It's also hard to put strain in a 
+framework with friction, for example. 'Physically realistic' modeling has to 
+make a lot of assumptions and use heavy duty equipment (finite elements, for 
+example) to incorporate strain.
 
-There is also a separate issue with elastic rebound theory: It's not very easy 
-to tell whether all the accumulated shear strain was released in an earthquake 
-or not. What kind of measurements would tell us this?
+There is also a separate issue with elastic rebound: It's not very easy to tell 
+whether all the accumulated shear strain was released in an earthquake or not. 
+What kind of measurements would tell us this?
 
@@
 decently described by Mohr-Coulomb models, and at, say, 10 km depth, the 
-confining pressure is almost 300 MPa. With reasonable static friction 
-coefficients (say 0.3 to 0.6) and typical pore fluid pressure,
+confining pressure is almost 300 MPa. With a reasonable rock density (2700 
+kg/m$^3$), coefficient of static friction (say 0.5), and pore fluid pressure 
+(say 0.3 times lithostatic pressure), the shear stress at failure is 94.5 MPa. 
+Stress drops are generally on the order of 0.1-10 MPa (see Peter Shearer's work 
+on Brune-type stress drop estimates, for example). So if less than 10% of shear 
+stress is relieved in an earthquake, what are the implications for the elastic 
+rebound hypothesis?
+
+My take on this is that elastic rebound is a great way to describe the 
+phenomenology of earthquakes to a non-geologist. Scientifically, it was an idea 
+of absolute genius in 1910, but it isn't a thorough or mechanically sound 
+framework for earthquake science a century later. Stress-based frameworks are 
+much better suited to both conceptual and quantitative treatments.
+
+*Changes: None.*
 
@@
 
-Isn't this implicit in the
+Good idea.  I added brief definitions of the scale, shape and location 
+parameters to an earlier paragraph where the terms are first written.
+
+*Changes: definitions added.*
 
@@
 distribution for magnitude-frequency distributions by Yan Kagan (and perhaps 
-others); this is similar in concept but not close enough to use here.
+others); this is not a similar-enough use to include here.
 
-(Moderately) interestingly, I gather that Pareto/power-law distributions often 
-form in environments characterized by exponential growth and aggregation 
-(population, money, etc.), while the somewhat similar and discussed 
-hyperexponential distribution can result from the sum of possibly-independent 
-random Poisson processes, though both give rise to final distributions with 
-similar tails.
-
-*No changes.*
+*Changes: None.*
 
@@
 
-*No changes.*
+*Changes: None.*
 
@@
 fault with a slip rate of 3 mm/yr and a per-event slip of 10 m, you scale the 
-slip by 10 and the slip rate by 3). It's not accidental.
+slip by 10 and the slip rate by 3). It's not accidental that I have chosen 
+these numbers, and I have explicitly described how to do the scaling.
 
-*No changes.*
+*Changes: None.*
 
@@
 > COV.
+
+Sorry to disappoint.
+
+A comparison between COV and the slip-rate variance at any time is actually 
+done in the study (it is in fact the core of the study)--I vary the COV of a 
+single distribution (the lognormal distribution) while changing nothing else, 
+then I keep the lognormal distribution with a COV of 1 and compare that to a 
+very different distribution (the exponential distribution) with a COV of 1. The 
+results, which are described unambiguously in the study, show that the slip 
+rate variance changes with COV but isn't much affected by the shape of the 
+distribution (i.e. lognormal vs. exponential).
+
+There is no reason to pursue this farther here. The work that I have done here 
+has covered the (small) space of geologically reasonable distributions; there 
+aren't huge gaps left unexplored.
+
+There are two main families of distributions that are in consideration for 
+earthquake recurrence: Exponential distributions (possibly with modifications 
+such as a stretched or hyperexponential distribution) and lognormal-like 
+distributions (lognormal, Wiebull, BPT, etc., which are not distinguishable in 
+real paleoseismic datasets). I have compared both types of distributions for a 
+single COV and the differences are very minor. There are no alternate 
+distribution families in any consideration within seismology that are *more* 
+different than these two families.
+
+This work is not meant to offer a mathematical proof, which is why I gave the 
+soft 'appears to be related to...' statement instead of a more firm 'is 
+demonstrated to be caused by...' or 'is proven to be a function of...'. 
+Demonstrating some correlation or relation is enough to further the general 
+point, which is that more earthquake recurrence variability will lead to more 
+short-term slip rate variability (which seems self-evident in any case).
+
+Going farther would either mean doing many more numerical experiments (which do 
+not offer real proof and would just clog the paper) or invoking more 
+complicated mathematical tools such as stochastic calculus, which I don't 
+really know and am unwilling to teach myself for this paper.
+
+*Changes: None.*
 
@@
 versa. This is quite unlike the autocorrelation in the Puget Lowlands of
-Washington State, which is more like a network of generally similar faults
-instead of an isolated or dominant plate boundary fault like the SAF or the
-Alpine Fault as referenced in another comment on Kelvin Berryman's work, where a
-long recurrence interval will more likely be followed by another long recurrence
-interval. I don't know what this means, or whether it's all noise, but it's
-intriguing to me and I put this bit in the paper in hopes of catching the
-attention of others who may be looking for a project.]
+Washington State, which is a network of generally similar faults; here 
+autocorrelation is positive, so a long recurrence interval will more likely be 
+followed by another long recurrence interval. I don't know what this means, or 
+whether it's all noise, but it's intriguing to me and I put this bit in the 
+paper in hopes of catching the attention of others who may be looking for a 
+project.]
 
@@
 near zero) friction is part and parcel with complete stress drop in major
-earthquakes, which is supported by many studies [REFS]. Furthermore, there are a
-host of laboratory experiments which suggest that friction during slip decreases
-to very low values [MORE REFS].
+earthquakes, which is supported by many studies (e.g., Hardebeck, 2012; 
+Hasegawa et al., 2011) though I generally disbelieve them. Furthermore, there 
+are a host of laboratory experiments which suggest that friction during slip 
+decreases to very low values (e.g., Do Toro et al., 2004).
+
+*Changes: None.*
+
+## Review by Anonymous
+
+> [A1] P1L1:  I suggest to use “aleatoric variability” and “epistemic 
+> uncertainty”. That way there is only one “variability” and one “uncertainty”, 
+> which makes the language more clear. Please modify throughout the manuscript.
+
+This is a fine suggestion.
+
+*Changes: modified terminology.*
+
+> [A2] P1L9: I believe the mathematicians call it just CV and not COV. Maybe a 
+> good idea to stick to the prior naming convention.
+
+Also a good suggestion.
+
+*Changes: modified acronym).*
+
+> [A3] P1L10:  This statement “ ... is quite high”
+> ... is a bit too vague.  Better add numbers (COV values) here as well.
+
+I modified the sentence in question to state that the rates may vary by a 
+factor of 3 or on short (<5,000 year) timescales.
+
+*Changes: Numbers added.*
+
+> [A4] P1L23: Putting the “e.g.” at the end of a list of references seems 
+> unusual. Is this an accepted format for this journal? Please check and modify 
+> in necessary.
+
+It was a LaTeX error on my part, and has been fixed.
+
+*Changes: LaTeX fix.*
+
+> [A5] P1L1:  The connection to locking depth should be explained a bit more.  
+> Good to also provide a reference here.
+
+I added a phrase stating that the width of the zone affected by 
+earthquake-cycle strains is a function of the fault's locking depth, and added 
+refs to Savage and Burford (1973), the classic 
+locked-fault-above-creeping-fault reference, and Hetland and Hager 2006, which 
+demonstrates that this process can instead be the result of post-seismic 
+relaxation.
+
+*Changes: explanation and references added.*
+
+> [A6] P3L29: I find it troublesome to talk about periodic/regular occurrence 
+> just because CV is smaller than 1. That would be correct for CV = 0, as you 
+> also pointed out. Depending on CV value between 0 and 1, it might be better 
+> to talk about quasi-periodic, or quasi- random behavior.
+
+The referenced sentence states that in this paper, I'm not using the word 
+'periodic' to mean perfectly periodic behavior. In any case I added a wiggle 
+phrase stating that I mean 'quasi-periodic' but I won't change the word through 
+the whole paper because it would decrease readability a bit.
+
+*Changes: terminology explained.*
+
+> [A7] P4L26:  Using this distribution seems plausible.  It would however be 
+> really interesting to see other distributions explored –if possible, that 
+> would be a great addition to make the manuscript more complete.
+
+I have added another numerical experiment using an empirical slip distribution; 
+see the response to [06a].
+
+*Changes: numerical experiment w/ alternate slip distribution added.*
+
+> [A8] P5L17:  Here you describe qualitatively how more or less closely the 
+> different distributions align with the mean slip rate.  While doing this 
+> qualitatively is ok to first order, I suggest that you go one step further 
+> and compute some form of misfit function i.e., residual (simplest a L1 or L2 
+> norm).
+
+An L-norm of any sort isn't appropriate because the numbers here aren't 
+pairwise (observation, model) data. Instead, there is a single 'true' value at 
+any time $t$, which is invariantly 1, and then hundreds to thousands of 
+simulated values at each time $t$.
+
+Instead, because the true value is 1, all of the results shown in Figure 5 are 
+quantified percentiles of misfit. While I could calculate the CV as suggested 
+in comment [A3], this is pretty reductive and the numbers at any time $t$ can 
+be retrieved from Figure 5.
+
+*Changes: None.*
+
+> [A9] P6L2: stress drop doesn’t need to be “complete” –just has to be “the 
+> same” each time to get to the outcome you describe here. Maybe better 
+> rephrase accordingly.
+
+No, this isn't true. The outcome that I describe is a correlation between 
+loading time and displacement, not an invariance of either loading time or 
+displacement. Having 'the same' stress drop doesn't predict correlated loading 
+time/recurrence intervals and displacement.
+
+It's possible that this paragraph was too confusingly written for it to be 
+easily interpretable. I have made some minor changes to the sentence structure 
+for clarification
+
+*Changes: Paragraph edits.*

### 2018-10-14 manuscript/eq_cycle_slip_rates.tex
finished revisions

@@
 \begin{abstract}
-Because of the natural variability (aleatoric uncertainty) in earthquake
+Because of the natural (aleatoric) variability in earthquake
 recurrence intervals and coseismic displacements on a fault, cumulative
@@
 exponential distribution. The results demonstrate that the most
-important parameter is the coefficient of variation (\(COV\) = standard
+important parameter is the coefficient of variation (\(CV\) = standard
 deviation / mean) of the recurrence distributions rather than the shape
 of the distribution itself. Slip rate variability over short time scales
-(\textless{} 10,000 years or 10 mean earthquake cycles) is quite high,
+(\textless{} 5,000 years or 5 mean earthquake cycles) is quite high, varying by 
+    a factor of 3 or more from the mean,
 but decreases with time and is close to stable after
@@
 variability is higher for recurrence distributions with a higher
-\(COV\). The natural variability in the slip rate estimates compared to
+\(CV\). The natural variability in the slip rate estimates compared to
 the true value is then used to estimate the epistemic uncertainty in a
@@
 of a few mean earthquake cycles (as in a paleoseismic slip rate
-estimate), but decreases rapidly to a factor of 1-2 with \textgreater{}5
+estimate), but decreases rapidly to a factor of 1—2 with \textgreater{}5
 mean earthquake cycles (as in a neotectonic slip rate study). These
@@
 the measurement uncertainties on the offset and time quantities
-\citep[e.g.]{bird_uncertainties_2007,zechar_incorporating_2009}.
+\citep[e.g.,][]{bird_uncertainties_2007,zechar_incorporating_2009}.
 However, for slip rate estimates on active faults made from offset
 measurements near the fault trace (i.e., within a horizontal distance
-that is a small fraction of the fault's locking depth), the episodic
+that is a small fraction of the fault's locking depth, as the width of the zone 
+affected by earthquake-cycle strains is a function of locking depth 
+\citep[e.g.,][]{savage_geodetic_1973, hetland_interseismic_2006}), the episodic
 nature of surface displacement due to the earthquake cycle will
@@
 function of the number of cumulative earthquakes that have contributed
-to the measured offset. For older Quaternary markers that have
-experienced tens to hundreds of major earthquakes, the effects will be
-minor, and for bedrock geologic markers with kilometers of displacement,
-the earthquake cycle is likely not worth accounting for. However, due to
-progressive erosion of geologic markers and the challenge of dating many
-late Pliocene to early Quaternary units (which are too old for
-radiocarbon and many cosmogenic nuclide systems), geologists often have
-no choice but to choose late Pleistocene to Holocene markers to date,
-and these units may also be more desirable targets if the scientists are
-primarily concerned with estimating the contemporary slip rate on a
-fault with a slip rate that may vary over Quaternary timescales
-\citep[e.g.]{rittase_temporal_2014,zinke_highly_2018}. For slow-moving
-faults, the slip either long waiting to be released, or recently
-released, may represent a sizeable fraction of the measured fault
-offset.
+to the measured offset (plus any aseismic strain such as afterslip or creep).
+For older Quaternary markers that have experienced tens to hundreds of major
+earthquakes, the effects will be minor, and for bedrock geologic markers with
+kilometers of displacement, the earthquake cycle is likely not worth accounting
+for. However, due to progressive erosion of geologic markers and the challenge
+of dating many late Pliocene to early Quaternary units (which are too old for
+radiocarbon and many cosmogenic nuclide systems), geologists often have no
+choice but to choose late Pleistocene to Holocene markers to date. These
+units may also be more desirable targets if the scientists are primarily
+concerned with estimating the contemporary slip rate on a fault with a slip
+rate that may vary over Quaternary timescales
+\citep[e.g.,][]{rittase_temporal_2014,zinke_highly_2018}. For slow-moving 
+faults, the slip either long waiting to be released, or recently released, may 
+represent a sizeable fraction of the measured fault offset.
 
@@
 available, especially in the years after a major earthquake
-\citep[e.g.]{rizza_earthquake_2015}, and many others will discuss the
-potential effects if the data are not \citep[e.g.]{lifton_latest_2015}.
+\citep[e.g.,][]{rizza_earthquake_2015}, and many others will discuss the
+potential effects if the data are not \citep[e.g.,][]{lifton_latest_2015}.
 These workers may only consider the time since the last earthquake,
@@
 However, the recurrence intervals between successive earthquakes on any
-given fault segment have some natural variability (i.e. \emph{aleatoric
-uncertainty}); similarly, displacement at a measured point is not
-identical in each earthquake \citep[e.g.]{duross_holocene_2008}.
+given fault segment have some natural (i.e. \emph{aleatoric})
+variability; similarly, displacement at a measured point is not
+identical in each earthquake \citep[e.g.,][]{duross_holocene_2008}.
 Therefore, the measured slip rate may deviate from the time-averaged
@@
 from geologic reconstructions are similar enough to those measured over a few 
-years through GPS geodesy that sudden, transient accelerations and decelerations
-are unlikely (e.g., **DEMETS AND DIXON 1999**). As a consequence, some plate
-boundary faults are among the most regularly-rupturing faults known,
-particularly sections that are isolated from nearby faults
-\citep[e.g.]{berryman_major_2012} and therefore not affected by stress
-perturbations resulting from earthquakes on other faults. These stress changes
-may be 'static' coseismic instantaneous stresses in the elastic upper crust resulting from
-earthquake displacement, or analogous stress changes in the viscoelastic lower
-crust or upper mantle from the time-dependent relaxation of static stress
-perturbations, or 'dynamic' transient stress changes that accompany the passage
-of seismic waves from nearby or distant earthquakes. Regardless of the nature
+years through GPS geodesy that sudden, transient accelerations and 
+decelerations are unlikely \citep[e.g.,][]{demets_new_1999}. As a consequence, 
+plate boundary faults may have near-constant rates of loading from tectonic 
+stress. Many plate boundary faults are are among the most regularly-rupturing 
+faults known, particularly sections that are isolated from nearby faults 
+\citep[e.g.,][]{berryman_major_2012} and therefore not affected by stress 
+perturbations resulting from earthquakes on other faults.
 
-Intraplate faults
+These stress perturbations may be `static' coseismic instantaneous stresses in 
+the elastic upper crust resulting from earthquake displacement 
+\citep{king_static_1994}, or analogous post-seismic stress changes in the 
+viscoelastic lower crust or upper mantle from the time-dependent relaxation of 
+static stress perturbations \citep[e.g.,][]{chery_postseismic_2001}, or 
+`dynamic' transient stress changes that accompany the passage of seismic waves 
+from nearby or distant earthquakes \citep{gomberg_dynamic_2005}.  Additionally, 
+changes in pore fluid pressure in a fault zone may increase or decrease the 
+required shear stress to initiate an earthquake 
+\citep{steacy_introduction_2005}. In contrast to isolated plate boundary 
+faults, intraplate faults or those on distributed plate boundaries may have 
+lower stress accumulation rates and the stress perturbations from activity on 
+other faults may be enough to significantly affect the timing of earthquakes on 
+a given fault \citep{gomberg_time-dependent_2005}.
 
-It is important to keep in mind that each earthquake in a sequence places
-points of rock on both sides of the fault that were not previously in contact.
-
-Though the physical mechanisms
-responsible and the statistical character of this natural variability
-remain under debate, its effects on the estimated slip rates may still
-be estimated given some common parameterizations.
+Though the physical mechanisms responsible and the statistical character of 
+this natural variability remain under debate, its effects on the estimated slip 
+rates may still be estimated given some common parameterizations.
 
@@
 study is geared towards providing useful heuristic bounds on the
-aleatoric and epistemic uncertainty of late Quaternary slip rate
+aleatoric variability and epistemic uncertainty of late Quaternary slip rate
 estimates for fault geologists, probabilistic seismic hazard modelers,
@@
 describe statistically, as it requires only one parameter (the mean rate
-parameter), which is a statistical scale parameter. The standard deviation of a
-large number of samples generated from an exponential distribution is equal to
-the mean.
+parameter), which is the inverse of the statistical scale parameter (the scale 
+parameter of a distribution determines the dispersion of the distribution, 
+while the shape parameter determines the shape or form of the distribution; a 
+parameter that shifts the distribution along the $x$-axis is called a location 
+parameter).  The standard deviation of a large number of samples generated from 
+an exponential distribution is equal to the mean.
 
@@
 done with a parameter known as the coefficient of variation, or
-\(COV = \sigma / \mu\), where \(\sigma\) is the standard deviation of
+\(CV = \sigma / \mu\), where \(\sigma\) is the standard deviation of
 the recurrence intervals, and \(\mu\) is the mean recurrence interval.
@@
 Periodic earthquakes are those that occur more regularly than random,
-and have a \(COV < 1\) (i.e. \(\sigma < \mu\)). These may be generated
+and have a \(CV < 1\) (i.e. \(\sigma < \mu\)). These may be generated
 by any of the time-dependent distributions described above with suitable
@@
 term `periodic' does not mean \emph{perfectly} repeating as it might in
-the physics or mathematics literature.)
+the physics or mathematics literature; the behavior referred to is most 
+accurately termed `quasi-periodic', but that term will not be used in the 
+interests of conciseness.)
 
 Unclustered earthquakes occur as regularly as random, and have a
-\(COV=1\). These may be generated by the exponential distribution (which
+\(CV=1\). These may be generated by the exponential distribution (which
 can generate no other), or by any of the time-dependent distributions as
@@
 Clustered earthquake sequences have sets of very tightly spaced
-earthquakes that are widely separated, and have a \(COV>1\). These may
+earthquakes that are widely separated, and have a \(CV>1\). These may
 be generated from a hyperexponential distribution, which is the sum of
@@
 periodic\\
-\citep[e.g.]{berryman_major_2012,scharer_quasi-periodic_2010}.
+\citep[e.g.,][]{berryman_major_2012,scharer_quasi-periodic_2010}.
 Conversely, many intraplate faults with low slip rates appear to show
 clustered earthquakes separated by long intervals of seismic quiescence
-\citep[e.g.]{clark_long-term_2012}. However, one can find examples of
+\citep[e.g.,][]{clark_long-term_2012}. However, one can find examples of
 studies indicating the opposite conclusions
@@
   distribution with a mean recurrence interval \(\mu\) = 1000 years, and
-  a standard deviation \(\sigma\) = 500 years, and a \(COV\) = 0.5.
+  a standard deviation \(\sigma\) = 500 years, and a \(CV\) = 0.5.
 \item
@@
   lognormal distribution with a mean recurrence interval \(\mu\) = 1000
-  years, a standard deviation \(\sigma\) = 1000 years, and a \(COV\) =
+  years, a standard deviation \(\sigma\) = 1000 years, and a \(CV\) =
   1.0.
@@
   lognormal distribution with a mean recurrence interval \(\mu\) = 1000
-  years, a standard deviation \(\sigma\) = 2000 years, and a \(COV\) =
+  years, a standard deviation \(\sigma\) = 2000 years, and a \(CV\) =
   2.0.
@@
   1000 years, a standard deviation \(\sigma\) = 1000 years, and a
-  \(COV\) = 1.0.
+  \(CV\) = 1.0.
 \end{enumerate}
@@
 behaviors with a compact and tractable number of simulations, and
-particularly to explore how changes in \(COV\) as well as the shape of
+particularly to explore how changes in \(CV\) as well as the shape of
 the distribution impact slip rate estimates.
@@
 some variability. This is representative of behavior observed in many
-studies \citep[e.g.]{zielke_slip_2010, klinger_characteristic_2011,
+studies \citep[e.g.,][]{zielke_slip_2010, klinger_characteristic_2011,
 zielke_earthquake_2018}. Taken together, the mean slip of 1 m and the
@@
 interested--can be easily given different shape and scale values to modify the
-\(COV\) or change the mean slip rate in the modeling code used in this paper.
+\(CV\) or change the mean slip rate in the modeling code used in this paper.
 
 However, it is not necessarily the most accurate representation of earthquake
-slip variability. **BIASI AND WELDON (2006)** compiled field measurements of
+slip variability. \citet{biasi_estimating_2006} compiled field measurements of
 surface ruptures from 13 earthquakes. The resulting distribution (Figure S1) has
 some significant differences with the lognormal distribution used here, though
-the \(COV\) of 0.67 is quite close to the value used here. To test the
+the \(CV\) of 0.67 is quite close to the value used here. To test the
 sensitivity of the results given in this paper to the choice of slip
 distribution, the numerical simulations in presented in this work were run with
-the only change being the use of the empirical slip distribution from **BIASI
-AND WELDON (2006)**, and the results are given in the supplemental material
-(Figure S2, Table S1). Though there is more discussion in the supplements, the
-results are essentially identical to those presented below. As an additional
-experiment, the numerical simulations have been run using an invariant per-event
-displacement of 1 m. Though this is not a realistic scenario, it allows for a
-deconvolution of the effects of earthquake time stochasticity and earthquake
-displacement stochasticity. The results are shown in Figure S3 and Table S2, and
-discussed in the supplemental materials; there are noticeable differences in the
-results, but they are small and do not call into question the results and
-conclusions presented below.
+the only change being the use of the empirical slip distribution from 
+\citet{biasi_estimating_2006}, and the results are given in the supplemental 
+material (Figure S2, Table S1). Though there is more discussion in the 
+supplements, the results are essentially identical to those presented below. As 
+an additional experiment, the numerical simulations have been run using an 
+invariant per-event displacement of 1 m. Though this is not a realistic 
+scenario, it allows for a deconvolution of the effects of earthquake time 
+stochasticity and earthquake displacement stochasticity. The results are shown 
+in Figure S3 and Table S2, and discussed in the supplemental materials; there 
+are noticeable differences in the results, but they are quite small and do not 
+call into question the results and conclusions presented below.
 
@@
 displacements can diverge greatly from the mean. The magnitude of this
-divergence appears to be related to the \(COV\) of the recurrence
-interval distributions: The clustered series (\(COV\) = 2) has by far
+divergence appears to be related to the \(CV\) of the recurrence
+interval distributions: The clustered series (\(CV\) = 2) has by far
 the most divergence, both unclustered series (lognormal and exponential
-with \(COV\) = 1) behave qualitatively similarly, and the periodic
-series (\(COV\) = 0.5) tracks most closely with the mean. The
+with \(CV\) = 1) behave qualitatively similarly, and the periodic
+series (\(CV\) = 0.5) tracks most closely with the mean. The
 divergences from the mean are driven by successive closely-spaced
@@
 series, or how much correlation is present between recurrence intervals
-and subsequent displacements. The former is essentially unstudied,
+and subsequent displacements. Autocorrelation in recurrence interval sequences 
+is essentially unstudied,
 though on the basis of preliminary, unreviewed analysis
 \citep{styron_survival_2017}, I suspect that it is as important as
-\(COV\). With respect to the latter, the framework of elastic rebound
+\(CV\). With respect to the amount of slip per event and loading time, the 
+framework of elastic rebound
 theory in its most basic form should predict some correspondence between
@@
 stress or strain release in each earthquake\\
-\citep[e.g.]{matthews_brownian_2002,dicaprio_post-seismic_2008} or in
+\citep[e.g.,][]{matthews_brownian_2002,dicaprio_post-seismic_2008} or in
 any model where coseismic friction drops to zero, as this is
@@
 \emph{epistemic} uncertainty in a slip rate measurement results from the
-\emph{aleatoric} uncertainty (or natural variability) in earthquake
+\emph{aleatoric} (or natural) variability in earthquake
 recurrence. However, the previous results have focused on describing the
@@
 much insight into fault processes such as growth
-\citep[e.g.]{roberts_fault_2002} and interaction
-\citep[e.g.]{wallace_grouping_1987,dolan_long-range_2007}. Practically,
+\citep[e.g.,][]{roberts_fault_2002} and interaction
+\citep[e.g.,][]{wallace_grouping_1987,dolan_long-range_2007}. Practically,
 if an older (or longer-term) slip rate is quite different from the
@@
 \(\hat{R_1}\) and \(\hat{R_2}\) and their empirical distributions
-(reflecting the number of earthquakes as well as the \(COV\)s of the
+(reflecting the number of earthquakes as well as the \(CV\)s of the
 underlying earthquake recurrence distributions are known, then the null
@@
 
-This work seeks to evaluate the effect of natural variability (aleatoric
-uncertainty) in earthquake recurrence intervals on slip rate
+This work seeks to evaluate the effect of natural (aleatoric) variability in 
+earthquake recurrence intervals on slip rate
 measurements. The study simulates cumulative displacement during
@@
   The most important factor in controlling the variability in slip rate
-  estimates is the coefficient of variation (\(COV\)); the different
+  estimates is the coefficient of variation (\(CV\)); the different
   distributions themselves are relatively unimportant. Faults with
-  \emph{periodic} earthquakes, with \(COV < 1\), have less initial
+  \emph{periodic} earthquakes, with \(CV < 1\), have less initial
   variability, and stabilize rapidly. Faults with \emph{unclustered}
-  earthquakes, with \(COV = 1\), have more variability and stabilize
+  earthquakes, with \(CV = 1\), have more variability and stabilize
   more slowly. Faults with \emph{clustered} earthquakes,\\
-  with \(COV > 1\), have a great amount of initial variability and
+  with \(CV > 1\), have a great amount of initial variability and
   require a large number of earthquakes to stabilize.
@@
 \tophline
-distribution  &  $t$     &     5  &     25 &     50 &    75 &   95 \\
+distribution  &  $t$     &     5\%  &   25\% &    50\% &   75\% &  95\% \\
 \middlehline
-logn\_0.5 & 2531  &  0.51 &  0.79 &  1.13 &  1.77 &  4.37 \\
+lognormal (CV=0.5) & 2531  &  0.51 &  0.79 &  1.13 &  1.77 &  4.37 \\
       & 4843  &  0.61 &  0.84 &  1.08 &  1.46 &  2.44 \\
@@
       & 42,103 &  0.83 &  0.96 &  1.04 &  1.12 &  1.27 \\
-logn\_1 & 2531  &  0.42 &  0.71 &  1.14 &  2.25 &   $\infty$ \\
+lognormal (CV=1) & 2531  &  0.42 &  0.71 &  1.14 &  2.25 &   $\infty$ \\
       & 4843  &  0.51 &  0.75 &  1.07 &  1.67 &  5.62 \\
@@
       & 42,103 &  0.72 &  0.89 &  1.03 &  1.18 &  1.55 \\
-logn\_2 & 2531  &  0.35 &  0.68 &  1.35 &   $\infty$ &   $\infty$ \\
+lognormal (CV=2) & 2531  &  0.35 &  0.68 &  1.35 &   $\infty$ &   $\infty$ \\
       & 4843  &  0.43 &   0.7 &  1.16 &  2.86 &   $\infty$ \\
@@
       & 42,103 &   0.7 &  0.83 &  0.98 &  1.28 &  4.75 \\
-exp\_1 & 2531  &   0.4 &   0.7 &  1.17 &  2.39 &   $\infty$ \\
+exponential (CV=1) & 2531  &   0.4 &   0.7 &  1.17 &  2.39 &   $\infty$ \\
       & 4843  &  0.49 &  0.75 &  1.08 &  1.67 &  4.79 \\
@@
 \end{tabular}
-  \caption{Epistemic uncertainty table \label{ep_unc_table}}
+  \caption{Epistemic uncertainty tabl, showing the percentiles for the 
+    slip-rate variability (in mm yr$^{-1}$) at each time $t$ (years). The 
+    long-term mean slip rate is 1 mm yr$^{-1}$. \label{ep_unc_table}}
 \end{table}

### 2018-10-14 manuscript/compile_tex.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+pdflatex eq_cycle_slip_rates.tex
+bibtex eq_cycle_slip_rates
+pdflatex eq_cycle_slip_rates.tex
+pdflatex eq_cycle_slip_rates.tex
+
+rm -f *.aux *.bbl *.log *.blg *.pagelist

### 2018-11-03 manuscript/reviews/response_letter.md
updates to response following third review

@@
 want this paper to be a simple, easy-to-digest paper and I think that a long 
-and necessarily unsatisfying (as we don't have answers yet) discussion on the 
+and necessarily unsatisfying discussion (as we don't have answers yet) on the 
 mechanisms behind recurrence variability will be an obstacle, and many readers 
@@
 *Changes: Paragraph edits.*
+
+## Review by Anonymous #2
+
+> [B1] Uncertainty in estimates of earthquake recurrence and fault slip rate are
+> important parameters, pursued by conscientious investigators of seismic
+> hazard. The author perceives a lack of statistical support, and offers in
+> this paper “insights” in variously tones of “friend of the practitioner” and
+> “trust me, I’m the numerate one here”. Neither is convincing.
+
+Someone's grumpy!
+
+Since this review seems to aim to shame rather than help, the number of changes
+are few.
+
+> [B2a] A couple of omissions in this paper are particularly striking.
+
+> [B2b] First, how do we have a paper addressed to “variability” in fault
+> slip rate, addressing particularly the problem of small samples, without
+> mentioning the methods of estimation for censored samples? There is an
+> extensive statistical literature to estimate parameters and uncertainties and
+> numerous recent papers applying it in paleoseismic contexts. This literature
+> provides real quantitative methods to deal with the open intervals, long or
+> short, that affect the geologist’s estimate of fault slip rate and recurrence
+> estimation. These are real equations, with real uncertainties. One would look
+> in vain in this paper for anything of similar substance.
+
+Open intervals and censoring are of course well trodden ground in the
+application of statistics to any field.
+
+However, much of the variability that this paper is concerned with is present
+within the closed intervals themselves. Open intervals
+
+I don't disagree that more substance is offered by methods developed through the
+decades by many researchers than I can offer in one short paper. Sorry!
+
+> [B2c] Second, pages of this paper could be replaced (and improved) by a
+> presentation and discussion of the properties of the standard error. E.g.,
+> given an estimate a sample- based estimate of the mean, how far might the
+> population (or true) mean be from the estimate? S.E. is estimated by the
+> sample standard deviation divided by the square root of the number of
+> samples. So, of course, estimates from small samples from a fuzzy log normal
+> converge more slowly than from a well-defined (quasi-periodic) lognormal.
+> Instead of a small equation (SE=s/sqrt(n)), our paper back-calculates the
+> result using 2 million years of samples, and presents the results like a new
+> discovery . And again, with little by way of meaningful uncertainties (e.g.,
+> p1, lines 11- 13, 14- 16).
+
+The standard error of the mean is irrelevant because slip rates are not
+generally derived by taking the mean of independently-measured per-earthquake
+offsets and recurrence intervals. They are typically derived by dividing net
+offset over net time (including or excluding open intervals as one wishes in
+paleoseismology, but never in neotectonics). Furthermore, with neotectonic slip
+rate estimates, $n$ is not known.
+
+It would be different if this paper was at all concerned with the calculation of
+recurrence intervals, per-event displacements, or other parameters of individual
+earthquakes where a number of samples
+
+> [B3] 1 L 9,10: We read that the most important parameter is the coefficient of
+>  variation. First, this equation is the arithmetic coefficient of variation,
+>  and not the CV for a lognormal distribution. The CV of a lognormal does not
+>  depend on the mean.
+
+The reviewer is really confused about the lognormal distribution (see [B6] and
+[B7] where the reviewer continues to make mistakes and level false accusations
+at me).
+
+I was quite worried that I made a mistake after reading this, so I re-checked
+my math, my code and re-read a range of references (e.g., *Statistical
+Distributions* by Evans, Hastings and Peacock, 1993; *Confidence Intervals
+for the Coefficent of Variation for the Normal and Lognormal Distributions*,
+Koopmans, Owen, and Rosenblatt, *Biometrika*, v. 51, no. 1/2, 1964, as well
+as a lot of internet pages). Nowhere do I find alternative definitions of the
+$CV$ for a lognormal distribution, except in the Wikipedia page for
+Coefficient of Variation where two conflicting definitions for the Geometric
+Coefficient of Variation are given. Both of these involve estimating the $CV$
+from samples rather than as an algebraic transformation of the equations for
+a lognormal distribution. The first definition is mathematically the same as
+the standard definition; the difference lies in whether the estimation
+procedure is done on the raw data or log-transformed data. The second is
+simply the square of the first.
+
+Just to reiterate, the arithmetic coefficient of variation for the lognormal
+distribution is the standard definition.
+
+The reviewer's confusion may stem from the fact that there are two equations
+for the coefficient of variation that get regularly passed around. However, as I
+demonstrate in the response to [B7c], they are mathematically equivalent. The
+simple definition, which I use in the paper, is $CV = \frac{\sigma}{\mu}$,
+where $\sigma$ and $\mu$ are the standard deviation and mean of the
+distribution. The other definition, $CV=\sqrt{e^{\phi^2}-1}$, is given in
+terms of the standard deviation $\phi$ of the natural log of the lognormal
+distribution.
+
+OK, this is pretty funny: All of the confusion that the reviewer has about the
+lognormal distribution appears to be because of a misreading of the
+Wikipedia pages for the lognormal distribution and for the coefficient of
+variation. I mentioned the geometric CV (non-)issue above.  The quip that the
+'CV of a lognormal does not depend on the mean' is... kind of true? As demonstrated in the response to [B7c], the standard deviation is a
+*function of the mean* in that the mean is a coefficient of the standard
+deviation. Therefore $CV = \sigma / \mu = \mu \Alpha /\mu$ where $\Alpha$ is
+another mathematical expression (see [B7c]). So, yes, technically it is true but
+only because it's on the top and the bottom of the fraction; furthermore, you
+can't derive the expression $A$ without knowing both $\mu$ and $\sigma$.
+Regardless, the quip gives all outward appearance of being a slight modification
+of the sentence "Contrary to the arithmetic standard deviation, the arithmetic
+coefficient of variation is independent of the arithmetic mean" which is below
+the equation for the $CV$ of a lognormal on the Wikipedia page.  It is not a
+statement that makes any practical sense unless one is beginning with
+log-transformed data or distributions.
+
+So to conclude, my math, code and analysis are sound in terms of the use of the
+$CV$.
+
+*No changes.*
+
+> [B4] We could stop here, but a central flaw in the paper is exposed – nothing
+> in this paper addresses how to obtain this most important parameter. If
+> attempted, the essential emptiness of a 2,000,000 year sample would emerge. No
+> real data set in paleoseismology resolves the mean and standard deviation to
+> better than maybe 50%. Typical sites do well to resolve it to a factor of 2.
+> P.3, line 22- 23 reflect this reality.
+
+"the essential emptiness"... I'm surprised, and perhaps a bit concerned, to
+find that I have tapped into the existential despair of Anonymous Reviewer #2.
+
+This study investigates how slip rates measured as *offset / time* of an offset
+marker are influenced by earthquake cycle variation. It is not a paper that
+discusses the finer points of calculating the mean and standard deviation of
+earthquake recurrence intervals or displacements from paleoseismology. Indeed,
+given that "There is an extensive literature..." [B2b] on this topic, why should
+I rehash it?
+
+I fully agree that a typical paleoseismological study cannot resolve the mean
+and standard deviation with a high degree of accuracy and I never stated
+otherwise. Partly for this reason, by using CVs of 0.5, 1, and 2, I cover a
+pretty wide range of values that can give some idea of the expected behavior of
+the system under
+
+> [B5a] It is not  obvious that the author has material experience words
+> “aleatoric” and “epistemic”. Line 1, “aleatoric uncertainty” is a
+> contradiction in terms.
+
+It's really not a contradiciton in terms. 'Aleatory' does not mean 'the absence
+of uncertainty', which it would have to in order to contradict the term
+'uncertainty'.
+
+Nonetheless
+
+O'Hagan, T. (2004). Dicing with the unknown. *Significance*, 1(3), 132-133.
+
+> [B5b] Bird, Zechar and Frankel all know better than to use the method the
+> author alleges in lines 21 and 22 to arrive at epistemic uncertainty in slip
+> rate.  They would more likely consider the allegation a misreading of their
+> work.
+
+This is a very perplexing comment. The line in question states, "The
+uncertainty in the resulting slip estimate is typically treated as epistemic,
+and quantified through the propagation of the measurement uncertainties on
+the offset and time quantities (e.g., Bird, 2007; Zechar and Frankel, 2009)."
+Both of the papers referenced are focused on estimating fault slip rates (or
+*long-term offset rates* in Bird's terminology) that incorporate the measurement
+uncertainty in both age and offset measurements. Both of their methods are based
+in a convolution of the probability distributions for age and offset of the
+marker features (i.e., the data with their uncertainties).
+
+> [B5c] More broadly, the lack of care in writing makes one wonder how to
+> understand this paper.
+
+Sorry!
+
+> [B5d] p2, L5:  A perturbation in slip rate would mean it was
+> slipping at rate X, then changes to Y.
+
+Yes.  This is both the intended and the obvious interpretation of the sentence.
+
+*No changes.*
+
+> [B5e] p.2, L14-17 have careful paleoseismologists doing reasonable things in
+> one sentence, then imply they would make plainly rookie mistakes in the next.
+
+The paragraph in question states that the open interval since the last
+earthquake is often taken into consideration, however the variability in closed
+intervals is not nearly as often taken into consideration--this is why I have
+written the paper. Should I not state that paleoseismologists and
+neotectonicists who consider the effects of the terminal open interval are
+careful? I don't feel like being less generous. Should I not point out that
+variability within the closed intervals is important and rarely considered?  No,
+I think that I should.
+
+*No changes.*
+
+> [B5f] From here these read like inexperienced generalizations.
+
+Sorry!
+
+*No changes.*
+
+> [B6a] p.4, L15-23: The descriptions of the lognormal variables here give one
+> pause. First, log-normal parameters do not have units.
+
+This comment is completely off the mark. The shape and scale parameters of a
+lognormal distribution are indeed unitless, as they are the mean and standard
+deviation of $\ln L$, where $L$ is a lognormal distribution (this transforms the
+lognormal distribution into a normal distribution). However, the mean and
+standard deviation of the actual lognormal distribution $L$ are in the units of
+the distribution itself.  Nowhere in the manuscript did I state that $\mu$ and
+$\sigma$, which are clearly defined as the mean and standard deviation of any
+distribution in question, are the shape and scale parameters of $L$.  They are
+simply the mean and standard deviation of it.
+
+I have added a footnote clarifying this and giving the derivation of the shape
+and scale parameters from the mean and standard deviation.
+
+*Changes: Footnote with explanation and equations added.*
+
+> [B7b] Second, the mean recurrence interval is not the location parameter of
+> a log normal. This is just wrong.
+
+Nowhere in the manuscript did I state this. In fact, the word 'location' is not
+present anywhere in the submitted manuscript (it has been added to an expanded
+discussion of probability distributions and their parameters in the revised
+manuscript as described in the response to Comment [O14], though I still do not
+make the mistake that I have been accused of).
+
+Furthermore, the lognormal distribution doesn't have a location parameter--it is
+a 2-parameter distribution specified by shape and scale parameters, as stated in
+P. 3, L. 24 of the submitted manuscript.
+
+*No changes.*
+
+> [B7c] Third, if one uses the CV equation for the lognormal distribution (e.g.,
+> https://en.wikipedia.org/wiki/Log-normal_distribution#Arithmetic_coefficient_of_variation),
+> the CV will not match the COV alleged here. Given that the study depends on
+> these distributions, we can’t really use subsequent conclusions.
+
+This is false. The equations are mathematically equivalent.
+
+The mean $\mu$ of a lognormal distribution $L$ is often given in terms of the
+mean $m$ and standard deviation $\phi$ of the log-transformed distribution 
+$\ln L$, which is a normal distribution. The equation is:
+
+$$\mu = e^{m + \frac{1}{2} \phi^2 }$$
+
+and similarly, the standard deviation $\sigma$ of $L$ may be given in terms of
+$m$ and $\phi$:
+
+$$\sigma = e^{m + \frac{1}{2} \phi^2} \sqrt{ e^{\phi^2} -1}\,,$$
+
+which is clearly equivalent to
+
+$$ \sigma = \mu \sqrt{ e^{\phi^2}-1}\,.$$
+
+Therefore,
+
+$$ CV = \frac{\sigma}{\mu} = \frac{\mu \sqrt{e^{\phi^2}-1}}{\mu} =
+\sqrt{e^{\phi^2}-1}$$
+
+which is the equation for the $CV$ of a lognormal distribution given in any
+source including the Wikipedia page.  Please note that in most sources, the
+symbols '$\mu$' and $\sigma$' are what I have here defined as '$m$' and '$\phi$'
+because I use the symbols $\mu$ and $\sigma$ to represent the mean and standard
+deviation of $L$, consistent with the manuscript where the same symbols are used
+equivalently across distributions.
+
+The methods and results of this study are consistent and this alleged error did
+not occur.
+
+*No changes.*
+
+> [B7] p.6, L20.  If the number of samples is really n = N – t + 1, the samples
+> are correlated by virtue of the overlap in the windows.  No accounting has
+> been made of the correlation structure.
+
+There is certainly a lot of autocorrelation of the samples if one were to
+look at them sequentially. Nonetheless, even if sequential samples are
+correlated, given enough samples, the final distribution of the samples won't
+depend on the autocorrelation in the samples (this principle underlies the
+family of powerful and commonly-used numerical methods based on Markov Chain
+Monte Carlo sampling, for instance). This is why I use 2,000,000 years of
+simulation: it ensures that the sampling is ample enough to not worry that
+the serially-correlated samples have not sufficiently sampled the
+distribution. I would have to use a much longer simulation (tens to hundreds
+of millions of years) otherwise.
+
+> [B8a] p.6, L26:  Starts a narrative of the consequences of the standard error,
+> as though the standard error was never invented.
+
+Again, the standard error is not applicable here because slip rates estimations
+as used in this study are not made by taking the mean and standard deviation of
+sets of measurements:  As laid out in
+
+> [B8b] The fuzzy, back-of-the-envelope estimates start to get thick here. Real
+> uncertainty estimates would serve better.
+
+This is kept fuzzy because it's not possible to know the recurrence or offset distribtions
+
+> [B9a] p.7, L1-3: Two observations: First, as written, the practicing
+> geologist is being asked to believe that 60 earthquake cycles have passed
+> with zero displacement. I can guess what was intended, but should not have to.
+
+Actually, the naive reading of the sentence, that 60 *mean* earthquake cycles
+can occur without displacement on a fault, is exactly what is intended.
+
+The practicing geologist is probably familiar with intraplate faults in
+environments with low strain rates.
+
+> [B9b] Second, what probability is associated with this 60-cycle thing?  I ask
+> because practicing hazard geologists have to make estimates, and give weights
+> to extreme events.  What is the probability of 60 cycles, a CV of 2.0, ...?
+> Hard to imagine that the author  has thought much about what these results
+> would mean or how to use them if they were true.
+
+I happen to be a practicing hazard geologist who is responsible for the
+implementation of fault sources for probabilistic seismic hazard analysis, so
+I have actually considered this
+
+There are two probabilities here:
+1. The probability of a fault with a CV of 2.0 going 60 mean earthquake cycles
+   without accruing much displacement. This one is pretty easy to answer just
+
+2. The probability of a fault having a CV of 2.0.  This is a goodFault 1 has a
+   slip rate of 1 mm/yr, and Fault 2 has a slip rate of 0.1 mm/yr. If I think of
+   nt of data.
+
+So what should the practicing geologist do here? Well, that all depends on the
+assumptions involved in the creation of the seismic source model. The first
+thing to do i  The first thing 
+to do i

### 2018-11-03 manuscript/eq_cycle_slip_rates.tex
updates to response following third review

@@
 periodic\\
-\citep[e.g.,][]{berryman_major_2012,scharer_quasi-periodic_2010}.
+\citep[e.g.,][]{berryman_major_2012, scharer_quasi-periodic_2010}.
 Conversely, many intraplate faults with low slip rates appear to show
@@
 studies indicating the opposite conclusions
-\citep{tuttle_earthquake_2002,grant_paleoseismic_1994}.
+\citep{tuttle_earthquake_2002, grant_paleoseismic_1994}.
 
@@
   distribution with a mean recurrence interval \(\mu\) = 1000 years, and
-  a standard deviation \(\sigma\) = 500 years, and a \(CV\) = 0.5.
+  a standard deviation \(\sigma\) = 500 years, and a \(CV\) =
+  0.5 \footnote{Please note that $\mu$ and $\sigma$ do not represent the scale
+  and shape parameters of the lognormal distribution, though these symbols are
+  commonly used to represent these parameters, as the scale and shape parameters
+  of a lognormal distribution $L$ are the standard deviation and mean of the
+  normal distribution $\ln L$. Instead, the shape parameter $sh$
+  and scale parameter $sc$ are derived from $\mu$ and $\sigma$ through the
+  relations $sh = \sqrt{1 + \sigma^2 / \mu^2}$ and 
+  $sc = \ln [ \mu / \sqrt{1 + \sigma^2 / \mu^2}]$.}
 \item

### 2018-11-26 manuscript/reviews/response_letter.md
done with revisions for now

@@
 ---
-title: Response to Reviewers, "The impact of earthquake cycle variability on 
+title: |
+  Response to Reviewers, "The impact of earthquake cycle variability on 
   neotectonic and paleoseismic slip rate estimates"
 author: 
-  name: Richard Styron
+- name: Richard Styron
 ---
@@
 document in the supplemental material. The most relevant figure is also shown
-here:
+here (Figure \ref{nsv_sre})
 
-![Figure \label{fig}](../figures/nsv_slip_rate_envelopes.pdf)
+![Slip rates with no displacement variability \label{nsv_sre}](../figures/nsv_slip_rate_envelopes.pdf)
 
@@
 are basically that these results are less smooth, but the total variance at any
-point in time (*x-axis*) is less. This is simply due to removing the
+point in time (*x*-axis) is less. This is simply due to removing the
 stochasticity from one of the two variables in the system. The relative
@@
 
-*Changes: None..*
+*Changes: None.*
 
@@
 after by a LaTeX bug that I hadn't diagnosed.
+
+*Changes: None.*
 
@@
 
+*Changes: None.*
+
 > [O20] Page 6, line 1. Zero friction at rupture arrest is very unrealistic, 
@@
 earthquakes, which is supported by many studies (e.g., Hardebeck, 2012;
-Hasegawa et al., 2011) though I generally disbelieve them. Furthermore, there 
-are a host of laboratory experiments which suggest that friction during slip 
-decreases to very low values (e.g., Do Toro et al., 2004).
+Hasegawa et al., 2011) though I am pretty suspicious of the results.
+Furthermore, there are a host of laboratory experiments which suggest that
+friction during slip decreases to very low values (e.g., Di Toro et al.,
+2004).
 
@@
 
-## Review by Anonymous
+## Review by Anonymous Revier #1
 
@@
 
-*Changes: modified acronym).*
+*Changes: modified acronym.*
 
@@
 I added a phrase stating that the width of the zone affected by
-earthquake-cycle strains is a function of the fault's locking depth, and added 
-refs to Savage and Burford (1973), the classic 
-locked-fault-above-creeping-fault reference, and Hetland and Hager 2006, which 
-demonstrates that this process can instead be the result of post-seismic 
-relaxation.
+earthquake-cycle strains is a function of the fault's locking depth, and
+added refs to Savage and Burford (1973), the classic
+locked-fault-above-creeping-fault reference, and Hetland and Hager (2006),
+which demonstrates that this process can instead be the result of
+post-seismic relaxation.
 
@@
 pairwise (observation, model) data. Instead, there is a single 'true' value at 
-any time $t$, which is invariantly 1, and then hundreds to thousands of 
+any time $t$, which is invariably 1, and then hundreds to thousands of 
 simulated values at each time $t$.
@@
 
-No, this isn't true. The outcome that I describe is a correlation between 
-loading time and displacement, not an invariance of either loading time or 
+No, this isn't true. The outcome that I describe is a *correlation* between 
+loading time and displacement, not an *invariance* of either loading time or 
 displacement. Having 'the same' stress drop doesn't predict correlated loading 
@@
 
-## Review by Anonymous #2
+## Review by Anonymous Reviewer #2
 
@@
 
-Open intervals and censoring are of course well trodden ground in the
-application of statistics to any field.
+The paper doesn't go into depth on the topics of open intervals and censored
+data because this paper does not use any samples of earthquake recurrence
+intervals or per-event displacements in the rate calculations. Though they
+are present behind the scenes, all of the numerical results in the paper are
+slip rates derived as the total offset over the total time (equation 1)--no
+information on how many earthquakes are present in that time window, or what
+the duration of open intervals may be, is used in the calculations. This is
+intended to simulate a neotectonic-type slip rate estimation (as written in page
+2, lines 24-25 of the original manuscript), where an offset geological marker
+layer is dated and the offset is measured. I have reinforced this by
+restating in the methods that neither the number of events or the length of
+the closed or open intervals are known.
 
-However, much of the variability that this paper is concerned with is present
-within the closed intervals themselves. Open intervals
+The decision to avoid discussing the topic of including censored data in
+parameter estimation is also warranted because the most robust method to
+incorporate open intervals are in maximum-likehood estimation, which is a
+complex topic to bring up, as are survival analysis and related methods. Since
+these topics are complimentary, not central, to this paper I don't see the
+benefit of adding a dense paragraph (much less any equations) for methods that
+aren't relevant.
 
@@
 decades by many researchers than I can offer in one short paper. Sorry!
+
+*Changes: brief reinforcement that this study uses neotectonic (age/offset) slip
+rate measurements, not those from multiple events as in a paleoseismic context.*
 
@@
 
-The standard error of the mean is irrelevant because slip rates are not
-generally derived by taking the mean of independently-measured per-earthquake
-offsets and recurrence intervals. They are typically derived by dividing net
-offset over net time (including or excluding open intervals as one wishes in
-paleoseismology, but never in neotectonics). Furthermore, with neotectonic slip
-rate estimates, $n$ is not known.
+The standard error does share the same basic decrease in dispersion with
+increasing number of events, but it is symmetric even though the distributions
+used here are not, and it does not account well for the systematic bias in short
+time windows that are observed in these results. The misfit between the
+symmetric and asymmetric approximations becomes particularly bad when estimating
+the 'epistemic' uncertainty as in Figure 6, which is the uncertainty that one
+should apply to a measurement to account for earthquake-cycle variability. I
+have calculated the standard error and added to this graph show in Figure
+\ref{se} below (not added to manuscript). The results do not support the use
+of the standard error here. The upper error envelope under-approximates the
+uncertainty at short time windows (due to the measurement bias), but then
+becomes acceptable. However, the lower error envelope stays at or below the
+5th percentile, though it should be the 1-standard deviation approximation.
 
+![Standard error (S.E.) plotted on top of Figure 6 from the manuscript. \label{se}](./epist_unc.pdf)
+
+The standard error is also not a great fit for neotectonic slip rate
+estimations: They are typically derived by dividing net offset over net time
+$n$ is not known, nor are the durations of the open intervals.
 It would be different if this paper was at all concerned with the calculation of
 recurrence intervals, per-event displacements, or other parameters of individual
-earthquakes where a number of samples
+earthquakes where a number of samples is used, but that is a different topic.
+
+Finally, meaningful uncertainties are given in Table 1 for all distributions at
+a wide range of times. It's not appropriate to reduce all of these to a single equation.
+
+*Changes: Short discussion of standard error added to 'slip rate calculations' section.*
 
@@
 Just to reiterate, the arithmetic coefficient of variation for the lognormal
-distribution is the standard definition.
+distribution is the standard definition. It is also equivalent to the geometric
+coefficient of variation for a theoretical longormal distribution (although
+not necessarily for sample-based estimates).
 
@@
 
-OK, this is pretty funny: All of the confusion that the reviewer has about the
-lognormal distribution appears to be because of a misreading of the
+OK, this is pretty funny: All of the confusion that the reviewer has about
+the lognormal distribution appears to be because of a misreading of the
 Wikipedia pages for the lognormal distribution and for the coefficient of
 variation. I mentioned the geometric CV (non-)issue above. The quip that the
-'CV of a lognormal does not depend on the mean' is... kind of true? As demonstrated in the response to [B7c], the standard deviation is a
-*function of the mean* in that the mean is a coefficient of the standard
-deviation. Therefore $CV = \sigma / \mu = \mu \Alpha /\mu$ where $\Alpha$ is
-another mathematical expression (see [B7c]). So, yes, technically it is true but
-only because it's on the top and the bottom of the fraction; furthermore, you
-can't derive the expression $A$ without knowing both $\mu$ and $\sigma$.
-Regardless, the quip gives all outward appearance of being a slight modification
-of the sentence "Contrary to the arithmetic standard deviation, the arithmetic
-coefficient of variation is independent of the arithmetic mean" which is below
-the equation for the $CV$ of a lognormal on the Wikipedia page.  It is not a
-statement that makes any practical sense unless one is beginning with
-log-transformed data or distributions.
+'CV of a lognormal does not depend on the mean' is... kind of true? As
+demonstrated in the response to [B7c], the standard deviation of a lognormal
+is a *function of the mean* in that the mean is a coefficient in the standard
+devation equation if it's expressed in the typical shape and scale
+parameters. Therefore $CV = \sigma / \mu = \mu A /\mu$ where $A$ is another
+mathematical expression (see [B7c]). So, yes, technically it is true but only
+because $\mu$ is on the top and the bottom of the fraction and therefore cancels
+out; furthermore, you can't derive the expression $A$ without knowing both
+$\mu$ and $\sigma$. Regardless, the quip gives all outward appearance of
+being a slight modification of the sentence "Contrary to the arithmetic
+standard deviation, the arithmetic coefficient of variation is independent of
+the arithmetic mean" which is below the equation for the $CV$ of a lognormal
+on the Wikipedia page. It is not a statement that makes any practical sense
+unless one is beginning with log-transformed data or distributions.
 
-So to conclude, my math, code and analysis are sound in terms of the use of the
-$CV$.
+So to conclude, my math, code and analysis are correct in terms of the use of
+the $CV$.
 
-*No changes.*
+*Changes: None.*
 
@@
 This study investigates how slip rates measured as *offset / time* of an offset
-marker are influenced by earthquake cycle variation. It is not a paper that
-discusses the finer points of calculating the mean and standard deviation of
-earthquake recurrence intervals or displacements from paleoseismology. Indeed,
-given that "There is an extensive literature..." [B2b] on this topic, why should
-I rehash it?
+marker are influenced by earthquake cycle variation. It is *not* a paper
+discussing methods or uncertainties invovlved in the calculation of the mean
+and standard deviation of earthquake recurrence intervals or displacements
+from paleoseismology. Indeed, if "There is an extensive statistical
+literature..." (comment [B2b]) on this topic, why should I rehash it?
 
@@
 pretty wide range of values that can give some idea of the expected behavior of
-the system under
+the almost any fault system if the basic recurrence distributions can be
+estimated or assumed.
+
+A discussion of 'guesstimating' the $CV$ is actually already present in the
+manuscript, starting on p.8, l. 31 in the original submission. However, I do
+not give an in-depth treatment of estimating the $CV$ on a fault--this topic
+would be suitable in a paper on interpreting field methods or a general
+overview of statistical paleoseismology, perhaps, but not in a numerical
+simulation paper that takes values of the $CV$ *a priori*.
+
+*Changes: None.*
 
@@
 
-Nonetheless
+In any case, the term 'aleatoric uncertainty' is commonly understood and used by
+statisticians in addition to us proles.  See for example O'Hagan, T. (2004). Dicing with the unknown. *Significance*, 1(3), 132-133.
 
-O'Hagan, T. (2004). Dicing with the unknown. *Significance*, 1(3), 132-133.
+*Changes: None.*
 
@@
 in a convolution of the probability distributions for age and offset of the
-marker features (i.e., the data with their uncertainties).
+marker features (i.e., the data with their uncertainties). Perhaps the reviewer
+takes 'the propagation of the measurement uncertainties...' to mean that it is
+done through adding them in quadrature to the final rate, but I clearly do not
+specify the methods used here and the reviewer may want to reign in the
+assumptions.
+
+*Changes: None.*
 
@@
 
+*No changes.*
+
 > [B5d] p2, L5:  A perturbation in slip rate would mean it was
@@
 Yes. This is both the intended and the obvious interpretation of the sentence.
+The slip rate at the fault trace varies because of variability in the frequency
+and magnitude of accumulating offsets.
 
@@
 variability within the closed intervals is important and rarely considered?  No,
-I think that I should.
+I think that I should. The first reference in this paragraph makes both the
+'reasonable' decision and then the 'rookie mistake'. I am not simply making this
+stuff up.
 
-*No changes.*
+*Changes: None.*
 
@@
 *No changes.*
+*Changes: None.*
 
@@
 deviation of $\ln L$, where $L$ is a lognormal distribution (this transforms the
-lognormal distribution into a normal distribution). However, the mean and
+lognormal distribution into a normal distribution).
+
+However, the mean and
 standard deviation of the actual lognormal distribution $L$ are in the units of
@@
 distribution in question, are the shape and scale parameters of $L$.  They are
-simply the mean and standard deviation of it.
+simply the mean and standard deviation of the distribution.
 
@@
 
+*Changes: None.*
+
 > [B8a] p.6, L26:  Starts a narrative of the consequences of the standard error,
@@
 
-Again, the standard error is not applicable here because slip rates estimations
-as used in this study are not made by taking the mean and standard deviation of
-sets of measurements:  As laid out in
+I have added a brief discussion of the standard error, which is an OK
+approximation of the results beyond 5-10 mean earthquake cycles (once the
+systematic bias in very short time windows has gone away), but the asymmetry of
+the results are not well approximated by the standard error. One could hack on
+it a bit by log-transforming the data, applying the standard error of the
+log-transformed distributions, and then anti-log transforming it, but that seems
+a bit much.
+
+*Changes: None.*
 
@@
 
-This is kept fuzzy because it's not possible to know the recurrence or offset distribtions
+The uncertainty estimates at a range of percentiles are available in Figure 5,
+and the uncertainties for a slip rate measurement are in Figure 6 and Table 1.
+
+*Changes: None.*
 
@@
 Actually, the naive reading of the sentence, that 60 *mean* earthquake cycles
-can occur without displacement on a fault, is exactly what is intended.
+passed without an earthquake on a fault in the simulation, is exactly what is
+intended.
 
-The practicing geologist is probably familiar with intraplate faults in
-environments with low strain rates.
+*No changes.*
 
@@
 implementation of fault sources for probabilistic seismic hazard analysis, so
-I have actually considered this
+I have actually considered this. It's very important when deciding whether to
+include faults that have very clear bedrock geomorphic signatures (say, from
+rapid deformation in the Miocene) and are favorably oriented for slip in the
+current stress regime, but lack evidence of Quaternary deformation. Are they to
+be included in the source models, or not?  I tend to add them, with low slip
+rates and high uncertainties, for just this reason. The analysis in this paper
+shows that very long recurrence intervals are possible on active faults.
 
-There are two probabilities here:
-1. The probability of a fault with a CV of 2.0 going 60 mean earthquake cycles
-   without accruing much displacement. This one is pretty easy to answer just
+There are two probabilities that are asked here:
+1. *The probability of a fault with a CV of 2.0 going 60 mean earthquake cycles
+   without accruing much displacement.*
 
-2. The probability of a fault having a CV of 2.0.  This is a goodFault 1 has a
-   slip rate of 1 mm/yr, and Fault 2 has a slip rate of 0.1 mm/yr. If I think of
-   nt of data.
+This is answered very easily using the equation for the lognormal probability distribution with a $CV$ of 2. The cumulative distribution function for
+this recurrence distribution at 60,000 years (60 mean earthquake cycles) is
+0.9994..., which means that there is a probability of about $p = 0.0006$ that
+any given recurrence interval will exceed 60,000 years, or 60 mean earthquake
+cycles. But the probability $P$ of a recurrence interval this long being present
+in a sequence of *n* earthquakes is $P=1 - (1-p)^n$, and for 2000 earthquakes
+(which is the approximate number in these simulations) the probability of a
+single recurrence interval exceeding 60 mean earthquake cycles is 0.1 or 10%.
+It becomes more likely than not that such a recurrence interval will be
+present in a sequence at about 12,500 earthquakes. Are you going to see this
+in a trench? Almost certainly not even if you were to have the appropriate
+sedimentation to preserve the signal. But that does not mean that it's an
+impossible or even improbable event in a long time series.
 
-So what should the practicing geologist do here? Well, that all depends on the
-assumptions involved in the creation of the seismic source model. The first
-thing to do i  The first thing 
-to do i
+*Changes*: I have added the following sentence to the paragraph under
+discussion: "It is highly unlikely that any given recurrence
+interval will be this long, but given thousands of earthquakes over millions
+of years, the chance of such an event occurring at least once is far more
+likely." I don't find it necessary to go through the probability calculations
+in the manuscript as I did here.
+
+2. *The probability of a fault having a CV of 2.0.*
+
+The best way to calculate this probability is basically to look at the
+frequency of such a CV occurring in some sample of faults. I don't have a big
+collection of faults with paleoseismic data to work with, and I'm not going
+to put one together to satisfy a reviewer, but I do calculate that the
+schematic, representative model for southeastern Australian intraplate faults
+by Clark et al., 2017 (cited in the paper) has a CV of 2. Unfortunately, Clark
+et al. cannot precisely date all of the earthquakes they infer so no real CV can
+be calculated from their data, but this is *their* best guess as to how the
+faults behave, based on their evidence for clusters of events with recurrence
+intervals of estimated mean 8,000 years separated by quiescent intervals of
+0.5-2 million years.
+
+Similarly, the USGS Quaternary Faults and Folds dataset has a number of faults
+that are thought to have last ruptured in the 'Middle and Late Quaternary' (<1.6
+Ma) but not in the 'Late Quaternary' (<130,000 years). Disregarding potential
+misclassifications of these faults, it seems reasonable to assume that a few of
+them may be experiencing very long recurrence intervals and may have a CV
+approaching 2. Many of these faults are along strike of fault segments with
+Holocene ruptures and there is no obvious geological difference between the
+segments with recent rupture and those without--the Steens fault zone in the
+lovely Steens rift of southeast Oregon is an example.
+
+*Changes: None.*

### 2018-11-26 manuscript/eq_cycle_slip_rates.tex
done with revisions for now

@@
 Please note that in the construction of the cumulative displacement
-histories, all samples are independent. This means that the duration of
-any recurrence interval does not depend on the duration of the previous
-or subsequent interval (in other words, there is no autocorrelation in
-these series); the same applies to the displacement samples.
-Furthermore, the magnitude of displacement is independent of the
-corresponding recurrence interval. It is currently unknown to what
-degree autocorrelation exists in real earthquake time and displacement
-series, or how much correlation is present between recurrence intervals
-and subsequent displacements. Autocorrelation in recurrence interval sequences 
-is essentially unstudied,
+histories, all samples are independent. This means that the duration of any
+recurrence interval does not depend on the duration of the previous or
+subsequent interval (in other words, there is no autocorrelation in these
+series); the same applies to the displacement samples. Furthermore, the
+magnitude of displacement is independent of the corresponding recurrence
+interval. It is currently unknown to what degree autocorrelation exists in
+real earthquake time and displacement series, or how much correlation is
+present between recurrence intervals and subsequent displacements.
+Autocorrelation in recurrence interval sequences is essentially unstudied,
 though on the basis of preliminary, unreviewed analysis
-\citep{styron_survival_2017}, I suspect that it is as important as
-\(CV\). With respect to the amount of slip per event and loading time, the 
-framework of elastic rebound
-theory in its most basic form should predict some correspondence between
-inter-event (loading) duration and slip magnitude, and this is included
-(implicitly or explicitly) in oscillator models incorporating complete
-stress or strain release in each earthquake\\
-\citep[e.g.,][]{matthews_brownian_2002,dicaprio_post-seismic_2008} or in
-any model where coseismic friction drops to zero, as this is
-functionally equivalent (because \(f^a = \tau^a_s / \tau^a_n\), where
-\(f^a\), \(\tau^a_s\) and \(\tau^a_n\) are respectively friction at
-rupture arrest, shear stress at rupture arrest, and effective normal
-stress at rupture arrest, zero friction implies zero shear stress, or
-complete stress drop). Given a reasonably constant loading rate,
-complete shear stress or strain release implies some proportionality
-between the loading time and displacement. Nonetheless, this
-correspondence is not found in the more extensive paleoseismic datasets,
-such as those by \citet{benedetti_earthquake_2013} (or the correlation
-may be negative as found by \citet{weldon_wrightwood_2004}), but the
-number of paleoseismic datasets of sufficient size and quantity to
-identify these effects with statistical significance are few indeed.
+\citep{styron_survival_2017}, I suspect that it is as important as \(CV\).
+With respect to the amount of slip per event and loading time, the framework
+of elastic rebound theory in its most basic form should predict some
+correspondence between inter-event (loading) duration and slip magnitude, and
+this is included (implicitly or explicitly) in oscillator models
+incorporating complete stress or strain release in each earthquake
+\citep[e.g.,][]{matthews_brownian_2002,dicaprio_post-seismic_2008} or in any
+model where coseismic friction drops to zero, as this is functionally
+equivalent (because \(f^a = \tau^a_s / \tau^a_n\), where \(f^a\),
+\(\tau^a_s\) and \(\tau^a_n\) are respectively friction at rupture arrest,
+shear stress at rupture arrest, and effective normal stress at rupture
+arrest, zero friction implies zero shear stress, or complete stress drop).
+Given a reasonably constant loading rate, complete shear stress or strain
+release implies some proportionality between the loading time and
+displacement. Nonetheless, this correspondence is not found in the more
+extensive paleoseismic datasets, such as those by
+\citet{benedetti_earthquake_2013} (or the correlation may be negative as
+found by \citet{weldon_wrightwood_2004}), but the number of paleoseismic
+datasets of sufficient size and cquality to identify these effects with
+statistical significance are few indeed.
 
@@
 along the displacement series. \(\hat{R}\) is calculated simply as
-
 \begin{equation}
@@
 \end{equation}
-
 where \(D_0\) is the cumulative displacement at the beginning of the
@@
 time window, and \(t\) is the length of the time window. The \(\hat{}\)
-symbol signifies an estimate rather than the true value \(R\). By
-sliding \(\hat{R}\) over the displacement series, a set of many samples
-of \(\hat{R}\) is generated, so that we may analyze the distribution.
-The number of samples is \(n = N - t + 1\), where \(N\) is the length of
-the total series (2,000,000 in this study).
+symbol signifies an estimate rather than the true value \(R\). This slip rate
+estimation method is intended to represent a neotectonic-style slip rate
+estimate in which the number of earthquakes that have contributed to the
+observed deformation is unknown, as are the durations of the open intervals that
+bound the time window (one of which preceeds deposition of the marker unit, and
+one is the time since the most recent earthquake and the measurement time).
+By sliding \(\hat{R}\) over the displacement series, a set of many samples of
+\(\hat{R}\) is generated, so that we may analyze the distribution. The number
+of samples is \(n = N - t + 1\), where \(N\) is the length of the total
+series (2,000,000 in this study).
 
-A major goal of this study is to provide an answer to the question,
-\emph{How long should slip rates be measured over in order to estimate a
-meaningful rate?}\\
-This question will be answered by looking at the distribution in
-\(\hat{R}\) as a function of \(t\). Fifty values of \(t\) from 500 years
-to 100,000 years, logarithmically spaced, are used. Note that given
-\(\mu\) of 1000 years, this translates to 0.5-100 mean numbers of
-earthquakes in the window.
+A major goal of this study is to provide an answer to the question, \emph{How
+long should slip rates be measured over in order to estimate a meaningful
+rate?} This question will be answered by looking at the distribution in
+\(\hat{R}\) as a function of \(t\). Fifty values of \(t\) from 500 years to
+100,000 years, logarithmically spaced, are used. Note that given \(\mu\) of
+1000 years, this translates to 0.5-100 mean numbers of earthquakes in the
+window.
 
 The results of these calculations are shown in Figure
-\ref{slip_rate_envelopes}. It is clear that the total variability in the
-estimated slip rates is initially quite high when \(t\) is short
-(\textless{}10,000 years or \textasciitilde{}10 earthquakes).
-Particularly when \(t\) is \textless{}5,000 years, the maximum rates are
-a factor of 3 or more greater than the true rate \(R\), but the median
-rates are lower than \(R\)---this means it is more likely that fewer
-earthquakes are captured in the time window than naively expected given
-the mean recurrence. As the median is lower than \(R\), most
-measurements over these short timescales will underestimate the mean
-rate, although not necessarily by much.
+\ref{slip_rate_envelopes} for up to 60 mean earthquake cycles. It is clear
+that the total variability in the estimated slip rates is initially quite
+high when \(t\) is short (\textless{}10,000 years or \textasciitilde{}10
+earthquakes). Particularly when \(t\) is \textless{}5,000 years, the maximum
+rates are a factor of 3 or more greater than the true rate \(R\), but the
+median rates are lower than \(R\)---this means it is more likely that fewer
+earthquakes are captured in the time window than naively expected given the
+mean recurrence, and that the time contained in the open intervals is a
+substantial fraction of the total time window. As the median is lower than
+\(R\), most measurements over these short timescales will underestimate the
+mean rate, although not necessarily by much.
 
@@
 the only exception here is that the lower bound of the clustered
-distribution can stay at zero for more than 60 mean earthquake cycles.
-For rate estimates longer than several tens of mean earthquake cycles,
-the variation decreases very slowly but progressively with increasing
-window length.
+distribution can stay at zero for more than 60 mean earthquake cycles. It is
+highly unlikely that any given recurrence interval will be this long, but
+given thousands of earthquakes over millions of years, the chance of such an
+event occurring at least once is far more likely. For rate estimates longer
+than several tens of mean earthquake cycles, the variation decreases very
+slowly but progressively with increasing window length.
+
+Note that with a measurement time exceeding 5-10 mean earthquake cycles, the
+standard error ($n / \sqrt{(\sigma_D^2 + \sigma_r^2)}$, where $\sigma_D$ is
+the standard deviation of displacement, and $\sigma_r$ is the standard
+deviation of the recurrence interval) is a reasonable approximation for the
+standard deviation of the variability in the slip rates due to
+earthquake-cycle stochasticity, and shows broadly similar decreasing
+variation with increasing earthquake cycles. However, the standard error is
+symmetrical, though the variability displayed here is asymmetrical due to the
+asymmetry of the recurrence interval and displacement distributions.
 
@@
 estimates, where there are rarely more than 5 events in any given
-trench. However, the median is less than 2 times \(\hat{R}\) following
-just 2-3 events, so the systematic bias is unlikely to be much greater
-than the measurement uncertainties on the age or offset of the events.
+trench, though this bias will be decreased if slip rate measurements are made
+from closed intervals only. However, the median is less than 2 times
+\(\hat{R}\) following just 2-3 events, so the systematic bias is unlikely to
+be much greater than the measurement uncertainties on the age or offset of
+the events.
 
@@
 and the number of earthquakes that have transpired are sufficently known
-to make the calculation, unless the fault has received sufficient
+to make the calculation, unless the fault has received in-depth
 paleoseismic and neotectonic study. As formal hypothesis testing may not

### 2018-11-26 scripts/slip_rate_fns.py

@@ -80,7 +80,7 @@ def make_cum_slip(rec_mean, rec_std, distribution, yrs=10000,
     
     cum_slip = np.cumsum(eq_slip)
     
-    return cum_slip
+    return eq_rec_times, cum_slip
 
 
 def moving_average_rate(cum_disp, window, scale='mm/yr'):

### 2018-12-10 manuscript/eq_cycle_slip_rates.tex
finished revised manuscript

@@
-\documentclass[se, manuscript]{copernicus}
+\documentclass[se]{copernicus}
 \usepackage{mathptmx}
@@
 \begin{abstract}
-Because of the natural (aleatoric) variability in earthquake
-recurrence intervals and coseismic displacements on a fault, cumulative
-slip on a fault does not increase linearly or perfectly step-wise with
-time; instead, some amount of variability in shorter-term slip rates
-results. Though this variability could greatly affect the accuracy of
-neotectonic (i.e., late Quaternary) and paleoseismic slip rate
-estimates, these effects have not been quantified. In this study,
-idealized faults with four different, representative earthquake
-recurrence distributions are created with equal mean recurrence
-intervals (1,000 years) and coseismic slip distributions, and the
-variability in slip rate measurements over 500 to 100,000 year
+Because of the natural (aleatoric) variability in earthquake recurrence
+intervals and coseismic displacements on a fault, cumulative slip on a fault
+does not increase linearly or perfectly step-wise with time; instead, some
+amount of variability in shorter-term slip rates results. Though this
+variability could greatly affect the accuracy of neotectonic (i.e., late
+Quaternary) and paleoseismic slip rate estimates, these effects have not
+been quantified. In this study, idealized faults with four different,
+representative earthquake recurrence distributions are created with equal
+mean recurrence intervals (1,000 years) and coseismic slip distributions,
+and the variability in slip rate estimates over 500 to 100,000 year
 measurement windows is calculated for all faults through Monte Carlo
-simulations. The recurrence distributions used are quasi-periodic,
-unclustered and clustered lognormal distributions, and an unclustered
-exponential distribution. The results demonstrate that the most
-important parameter is the coefficient of variation (\(CV\) = standard
-deviation / mean) of the recurrence distributions rather than the shape
-of the distribution itself. Slip rate variability over short time scales
-(\textless{} 5,000 years or 5 mean earthquake cycles) is quite high, varying by 
-    a factor of 3 or more from the mean,
-but decreases with time and is close to stable after
-\textasciitilde{}40,000 years (40 mean earthquake cycles). This
-variability is higher for recurrence distributions with a higher
-\(CV\). The natural variability in the slip rate estimates compared to
-the true value is then used to estimate the epistemic uncertainty in a
-single slip rate measurement (as one would make in a geological study)
-in the absence of any measurement uncertainty. This epistemic
-uncertainty is very high (a factor of 2 or more) for measurement windows
-of a few mean earthquake cycles (as in a paleoseismic slip rate
-estimate), but decreases rapidly to a factor of 1—2 with \textgreater{}5
-mean earthquake cycles (as in a neotectonic slip rate study). These
-uncertainties are independent of, and should be propagated with
-uncertainties in fault displacement and geochronologic measurements used
-to estimate slip rates. They may then aid in the comparison of slip
-rates from different methods or the evaluation of potential slip rate
-changes over time.
+simulations. Slip rates are calculated as net offset divided by elapsed
+time, as in a typical neotectonic study. The recurrence distributions used
+are quasi-periodic, unclustered and clustered lognormal distributions, and
+an unclustered exponential distribution. The results demonstrate that the
+most important parameter is the coefficient of variation (\(CV\) = standard
+deviation / mean) of the recurrence distributions rather than the shape of
+the distribution itself. Slip rate variability over short time scales
+(\textless{}5,000 years or 5 mean earthquake cycles) is quite high, varying
+by a factor of 3 or more from the mean, but decreases with time and is
+close to stable after \textasciitilde{}40,000 years (40 mean earthquake
+cycles). This variability is higher for recurrence distributions with a
+higher \(CV\). The natural variability in the slip rate estimates compared
+to the true value is then used to estimate the epistemic uncertainty in a
+single slip rate measurement (as one would make in a geological study) in
+the absence of any measurement uncertainty. This epistemic uncertainty is
+very high (a factor of 2 or more) for measurement windows of a few mean
+earthquake cycles (as in a paleoseismic slip rate estimate), but decreases
+rapidly to a factor of 1—2 with \textgreater{}5 mean earthquake cycles (as
+in a neotectonic slip rate study). These uncertainties are independent of,
+and should be propagated with, uncertainties in fault displacement and
+geochronologic measurements used to estimate slip rates. They may then aid
+in the comparison of slip rates from different methods or the evaluation of
+potential slip rate changes over time.
 \end{abstract}
@@
 
-Fault slip rates are generally estimated by dividing measurements of the
-offset of geologic marker features by the time over which that offset
-accumulated. The uncertainty in the resulting slip estimate is typically
-treated as \emph{epistemic}, and quantified through the propagation of
-the measurement uncertainties on the offset and time quantities
-\citep[e.g.,][]{bird_uncertainties_2007,zechar_incorporating_2009}.
-However, for slip rate estimates on active faults made from offset
-measurements near the fault trace (i.e., within a horizontal distance
-that is a small fraction of the fault's locking depth, as the width of the zone 
-affected by earthquake-cycle strains is a function of locking depth 
-\citep[e.g.,][]{savage_geodetic_1973, hetland_interseismic_2006}), the episodic
-nature of surface displacement due to the earthquake cycle will
-necessarily affect the results: If the measurements are taken
-immediately before an earthquake, the measured offset and resulting slip
-rate estimate will be lower than average, while if the measurements are
-taken immediately after an earthquake, the offset and rate will be
-higher.
+Fault slip rates are generally estimated by dividing measurements of the offset
+of geologic marker features by the time over which that offset accumulated (it
+is not currently possible to measure a slip rate directly, though the term
+`slip rate measurement' may be used to compare to a simulated or modeled
+value). The uncertainty in the resulting slip estimate is typically treated as
+\emph{epistemic}, and quantified through the propagation of the measurement
+uncertainties on the offset and time quantities
+\citep[e.g.,][]{bird_uncertainties_2007, zechar_incorporating_2009}.  However,
+for slip rate estimates on active faults made from offset measurements near the
+fault trace (i.e., within a horizontal distance that is a small fraction of the
+fault's locking depth, as the width of the zone affected by earthquake-cycle
+strains is a function of locking depth \citep[e.g.,][]{savage_geodetic_1973,
+hetland_interseismic_2006}), the episodic nature of surface displacement due to
+the fault's position in the earthquake cycle will necessarily affect the
+results: If the measurements are taken immediately before an earthquake, the
+measured offset and resulting slip rate estimate will be lower than average,
+while if the measurements are taken immediately after an earthquake, the offset
+and rate will be higher.
 
-The magnitude of the perturbation to the slip rate is, of course, a
+The magnitude of the perturbation to the slip rate estimate is, of course, a
 function of the number of cumulative earthquakes that have contributed
@@
 
-\section{Modeling the earthquake
-cycle}\label{modeling-the-earthquake-cycle}
+\section{Modeling the earthquake cycle}\label{modeling-the-earthquake-cycle}
 
@@
 describe statistically, as it requires only one parameter (the mean rate
-parameter), which is the inverse of the statistical scale parameter (the scale 
+parameter), which is the inverse of the statistical scale parameter. (The scale 
 parameter of a distribution determines the dispersion of the distribution, 
@@
 parameter that shifts the distribution along the $x$-axis is called a location 
-parameter).  The standard deviation of a large number of samples generated from 
+parameter.) The standard deviation of a large number of samples generated from 
 an exponential distribution is equal to the mean.
@@
 
-Clustered earthquake sequences have sets of very tightly spaced
-earthquakes that are widely separated, and have a \(CV>1\). These may
-be generated from a hyperexponential distribution, which is the sum of
+\begin{figure}[t]
+\includegraphics[width=8.3cm]{./figures/recurrence_dists.pdf}
+\caption{Earthquake recurrence distributions. `logn' = lognormal. `exp' = 
+  exponential. Colors for each distributions are the same in all figures. 
+  \label{eq_rec_dists}}
+\end{figure}
+
+Clustered earthquake sequences have sets of very tightly spaced earthquakes
+that are widely separated (Figure \ref{eq_spacing}, and have a \(CV>1\). These
+may be generated from a hyperexponential distribution, which is the sum of
 multiple exponentials with different means, or from the time-dependent
@@
 \citep[e.g.,][]{clark_long-term_2012}. However, one can find examples of
-studies indicating the opposite conclusions
+studies indicating the opposite conclusions, even from the same study areas
 \citep{tuttle_earthquake_2002, grant_paleoseismic_1994}.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{./figures/event_spacing.pdf}
+  \caption{Spacing of 15 simulated successive earthquakes from each recurrence 
+  distribution. Note that the gap between the last displayed earthquake and the 
+  right side of the plot does not represent a long recurrence interval. 
+  \label{eq_spacing}}
+\end{figure}
 
@@
   0.5 \footnote{Please note that $\mu$ and $\sigma$ do not represent the scale
-  and shape parameters of the lognormal distribution, though these symbols are
-  commonly used to represent these parameters, as the scale and shape parameters
-  of a lognormal distribution $L$ are the standard deviation and mean of the
-  normal distribution $\ln L$. Instead, the shape parameter $sh$
+  and shape parameters of the lognormal distribution in this work, though these
+  symbols are commonly used to represent these parameters elsewhere as the
+  scale and shape parametersof a lognormal distribution $L$ are the standard
+  deviation and mean of the normal distribution $\ln L$. Instead, the shape
+  parameter $sh$
   and scale parameter $sc$ are derived from $\mu$ and $\sigma$ through the
@@
 zielke_earthquake_2018}. Taken together, the mean slip of 1 m and the
-mean recurrence interval of 1000 years shared by each of the recurrence
+mean recurrence interval of 1,000 years shared by each of the recurrence
 interval distributions yields a mean slip rate of 1 mm/yr. This rate is
@@
 be generalized to faults with different parameters.
+
+\begin{figure}[t]
+\includegraphics[width=8.3cm]{./figures/slip_dist.pdf}
+\caption{Earthquake slip distribution. \label{slip_dist}}
+\end{figure}
 
@@
 randomly. Then, these are combined with an equal number of displacement
-samples drawn randomly from the earthquake slip distribution. Finally a
+samples drawn randomly from the earthquake slip distribution. Finally, a
 cumulative displacement history is created for each series from a
@@
 
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{./figures/disp_histories.pdf}
+  \caption{Simulated displacement histories for each of the recurrence 
+  distributions, and the `true' mean line at 1 mm/yr in black. \emph{A}: The 
+  first hundred thousand years. \emph{B}: The entire 2 million years. The 
+  histories are the same in both plots. \label{disp_histories}}
+\end{figure*}
+
 Please note that in the construction of the cumulative displacement
@@
 subsequent interval (in other words, there is no autocorrelation in these
-series); the same applies to the displacement samples. Furthermore, the
-magnitude of displacement is independent of the corresponding recurrence
-interval. It is currently unknown to what degree autocorrelation exists in
+series); the same applies to the displacement samples. It is currently unknown 
+to what degree autocorrelation exists in
 real earthquake time and displacement series, or how much correlation is
@@
 \citep{styron_survival_2017}, I suspect that it is as important as \(CV\).
-With respect to the amount of slip per event and loading time, the framework
-of elastic rebound theory in its most basic form should predict some
-correspondence between inter-event (loading) duration and slip magnitude, and
-this is included (implicitly or explicitly) in oscillator models
-incorporating complete stress or strain release in each earthquake
-\citep[e.g.,][]{matthews_brownian_2002,dicaprio_post-seismic_2008} or in any
-model where coseismic friction drops to zero, as this is functionally
-equivalent (because \(f^a = \tau^a_s / \tau^a_n\), where \(f^a\),
-\(\tau^a_s\) and \(\tau^a_n\) are respectively friction at rupture arrest,
-shear stress at rupture arrest, and effective normal stress at rupture
-arrest, zero friction implies zero shear stress, or complete stress drop).
-Given a reasonably constant loading rate, complete shear stress or strain
-release implies some proportionality between the loading time and
-displacement. Nonetheless, this correspondence is not found in the more
-extensive paleoseismic datasets, such as those by
-\citet{benedetti_earthquake_2013} (or the correlation may be negative as
-found by \citet{weldon_wrightwood_2004}), but the number of paleoseismic
-datasets of sufficient size and cquality to identify these effects with
-statistical significance are few indeed.
+
+Furthermore, the magnitude of displacement is independent of the corresponding
+recurrence interval. The framework of elastic rebound theory in its most basic
+form should predict some correspondence between inter-event (loading) duration
+and slip magnitude, and this is included (implicitly or explicitly) in
+oscillator models incorporating complete stress or strain release in each
+earthquake \citep[e.g.,][]{matthews_brownian_2002,dicaprio_post-seismic_2008}
+or in any model where coseismic friction drops to zero, as this is functionally
+equivalent (because \(f^a = \tau^a_s / \tau^a_n\), where \(f^a\), \(\tau^a_s\)
+and \(\tau^a_n\) are respectively friction at rupture arrest, shear stress at
+rupture arrest, and effective normal stress at rupture arrest, zero friction
+implies zero shear stress, or complete stress drop). Given a reasonably
+constant loading rate, complete shear stress or strain release implies some
+proportionality between the loading time and displacement. Nonetheless, this
+correspondence is not found in the more extensive paleoseismic datasets, such
+as those by \citet{benedetti_earthquake_2013} (or the correlation may be
+negative as found by \citet{weldon_wrightwood_2004}), but the number of
+paleoseismic datasets of sufficient size and cquality to identify these effects
+with statistical significance are few indeed.
 
@@
 100,000 years, logarithmically spaced, are used. Note that given \(\mu\) of
-1000 years, this translates to 0.5-100 mean numbers of earthquakes in the
+1000 years, this translates to 0.5--100 mean numbers of earthquakes in the
 window.
@@
 
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{./figures/slip_rate_envelopes.pdf}
+  \caption{Envelopes of estimated slip rates as a function of the mean number 
+  of earthquakes (or thousands of years) over which the slip rate was 
+  estimated. All slip rates have a true value of 1 mm/yr. \emph{A}: periodic 
+  distribution. \emph{B}: unclustered lognormal distribution. \emph{C}: 
+  Clustered lognormal distribution. \emph{C}: Unclustered exponential 
+  distribution. \label{slip_rate_envelopes}}
+\end{figure*}
+
 With longer \(t\), between 10,000--20,000 years (or 10--20 earthquakes)
@@
 
-Note that with a measurement time exceeding 5-10 mean earthquake cycles, the
+Note that with a measurement time exceeding 5--10 mean earthquake cycles, the
 standard error ($n / \sqrt{(\sigma_D^2 + \sigma_r^2)}$, where $\sigma_D$ is
@@
 
-\subsection{Interpreting measured
-rates}\label{interpreting-measured-rates}
+\subsection{Interpreting measured rates}\label{interpreting-measured-rates}
 
@@
 
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{./figures/epist_unc.pdf}
+  \caption{Epistemic uncertainty relative to the measured rate for each of the 
+  recurrence distributions, as a function of the mean number of earthquakes (or 
+  thousands of years) over which the slip rates were measured. \emph{A}: 
+  periodic distribution. \emph{B}: unclustered lognormal distribution. 
+  \emph{C}: Clustered lognormal distribution. \emph{C}: Unclustered exponential 
+  distribution.
+  \label{epist_unc}}
+\end{figure*}
+
 First, the variance in the distributions is quite large for the first
@@
 
-Second, the distributions are asymmetrical, especially the 5-95\%
+\begin{table}[t]
+\begin{tabular}{lllllll}
+\tophline
+distribution  &  $t$     &     5\%  &   25\% &    50\% &   75\% &  95\% \\
+\middlehline
+lognormal (CV=0.5) & 2531  &  0.51 &  0.79 &  1.13 &  1.77 &  4.37 \\
+      & 4843  &  0.61 &  0.84 &  1.08 &  1.46 &  2.44 \\
+      & 10,323 &   0.7 &  0.89 &  1.04 &  1.27 &  1.84 \\
+      & 42,103 &  0.83 &  0.96 &  1.04 &  1.12 &  1.27 \\
+lognormal (CV=1) & 2531  &  0.42 &  0.71 &  1.14 &  2.25 &   $\infty$ \\
+      & 4843  &  0.51 &  0.75 &  1.07 &  1.67 &  5.62 \\
+      & 10,323 &   0.6 &  0.82 &  1.04 &  1.38 &  2.55 \\
+      & 42,103 &  0.72 &  0.89 &  1.03 &  1.18 &  1.55 \\
+lognormal (CV=2) & 2531  &  0.35 &  0.68 &  1.35 &   $\infty$ &   $\infty$ \\
+      & 4843  &  0.43 &   0.7 &  1.16 &  2.86 &   $\infty$ \\
+      & 10,323 &  0.51 &  0.75 &  1.07 &  1.66 &   $\infty$ \\
+      & 42,103 &   0.7 &  0.83 &  0.98 &  1.28 &  4.75 \\
+exponential (CV=1) & 2531  &   0.4 &   0.7 &  1.17 &  2.39 &   $\infty$ \\
+      & 4843  &  0.49 &  0.75 &  1.08 &  1.67 &  4.79 \\
+      & 10,323 &  0.61 &   0.8 &  1.03 &  1.37 &  2.32 \\
+      & 42,103 &  0.76 &  0.89 &     1 &  1.15 &  1.43 \\
+\bottomhline
+\end{tabular}
+  \caption{Epistemic uncertainty tabl, showing the percentiles for the 
+    slip-rate variability (in mm yr$^{-1}$) at each time $t$ (years). The 
+    long-term mean slip rate is 1 mm yr$^{-1}$. \label{ep_unc_table}}
+\end{table}
+
+Second, the distributions are asymmetrical, especially the 5--95\%
 interval. The 95th percentile is generally several times as far from the
@@
 from closed intervals only. However, the median is less than 2 times
-\(\hat{R}\) following just 2-3 events, so the systematic bias is unlikely to
+\(\hat{R}\) following just 2--3 events, so the systematic bias is unlikely to
 be much greater than the measurement uncertainties on the age or offset of
@@
   mean earthquake cycles is very large, but begins to stabilize
-  following \textasciitilde{}10-20 earthquakes.
+  following \textasciitilde{}10--20 earthquakes.
 \item
@@
 
-\begin{figure}[t]
-\includegraphics[width=8.3cm]{./figures/recurrence_dists.pdf}
-\caption{Earthquake recurrence distributions. `logn' = lognormal. `exp' = 
-  exponential. Colors for each distributions are the same in all figures. 
-  \label{eq_rec_dists}}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{./figures/event_spacing.pdf}
-  \caption{Spacing of 15 simulated successive earthquakes from each recurrence 
-  distribution. Note that the gap between the last displayed earthquake and the 
-  right side of the plot does not represent a long recurrence interval. 
-  \label{eq_spacing}}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[t]
-\includegraphics[width=8.3cm]{./figures/slip_dist.pdf}
-\caption{Earthquake slip distribution. \label{slip_dist}}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{./figures/disp_histories.pdf}
-  \caption{Simulated displacement histories for each of the recurrence 
-  distributions, and the `true' mean line at 1 mm/yr in black. \emph{A}: The 
-  first hundred thousand years. \emph{B}: The entire 2 million years. The 
-  histories are the same in both plots. \label{disp_histories}}
-\end{figure*}
-
-\clearpage
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{./figures/slip_rate_envelopes.pdf}
-  \caption{Envelopes of estimated slip rates as a function of the mean number 
-  of earthquakes (or thousands of years) over which the slip rate was 
-  estimated. All slip rates have a true value of 1 mm/yr. \emph{A}: periodic 
-  distribution. \emph{B}: unclustered lognormal distribution. \emph{C}: 
-  Clustered lognormal distribution. \emph{C}: Unclustered exponential 
-  distribution. \label{slip_rate_envelopes}}
-\end{figure*}
-
-\clearpage
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{./figures/epist_unc.pdf}
-  \caption{Epistemic uncertainty relative to the measured rate for each of the 
-  recurrence distributions, as a function of the mean number of earthquakes (or 
-  thousands of years) over which the slip rates were measured. \emph{A}: 
-  periodic distribution. \emph{B}: unclustered lognormal distribution. 
-  \emph{C}: Clustered lognormal distribution. \emph{C}: Unclustered exponential 
-  distribution.
-  \label{epist_unc}}
-\end{figure*}
-
-\clearpage
-
-\begin{table}[t]
-\begin{tabular}{lllllll}
-\tophline
-distribution  &  $t$     &     5\%  &   25\% &    50\% &   75\% &  95\% \\
-\middlehline
-lognormal (CV=0.5) & 2531  &  0.51 &  0.79 &  1.13 &  1.77 &  4.37 \\
-      & 4843  &  0.61 &  0.84 &  1.08 &  1.46 &  2.44 \\
-      & 10,323 &   0.7 &  0.89 &  1.04 &  1.27 &  1.84 \\
-      & 42,103 &  0.83 &  0.96 &  1.04 &  1.12 &  1.27 \\
-lognormal (CV=1) & 2531  &  0.42 &  0.71 &  1.14 &  2.25 &   $\infty$ \\
-      & 4843  &  0.51 &  0.75 &  1.07 &  1.67 &  5.62 \\
-      & 10,323 &   0.6 &  0.82 &  1.04 &  1.38 &  2.55 \\
-      & 42,103 &  0.72 &  0.89 &  1.03 &  1.18 &  1.55 \\
-lognormal (CV=2) & 2531  &  0.35 &  0.68 &  1.35 &   $\infty$ &   $\infty$ \\
-      & 4843  &  0.43 &   0.7 &  1.16 &  2.86 &   $\infty$ \\
-      & 10,323 &  0.51 &  0.75 &  1.07 &  1.66 &   $\infty$ \\
-      & 42,103 &   0.7 &  0.83 &  0.98 &  1.28 &  4.75 \\
-exponential (CV=1) & 2531  &   0.4 &   0.7 &  1.17 &  2.39 &   $\infty$ \\
-      & 4843  &  0.49 &  0.75 &  1.08 &  1.67 &  4.79 \\
-      & 10,323 &  0.61 &   0.8 &  1.03 &  1.37 &  2.32 \\
-      & 42,103 &  0.76 &  0.89 &     1 &  1.15 &  1.43 \\
-\bottomhline
-\end{tabular}
-  \caption{Epistemic uncertainty tabl, showing the percentiles for the 
-    slip-rate variability (in mm yr$^{-1}$) at each time $t$ (years). The 
-    long-term mean slip rate is 1 mm yr$^{-1}$. \label{ep_unc_table}}
-\end{table}
-
 \end{document}
